---
title: "Chi-Square (2×2)"
subtitle: "Pearson's chi-square test of independence for two categorical variables"
order: 3
---

## Overview

This guide covers the 2×2 chi-square test of independence workflow in
**psych350lab**: running the analysis, creating contingency tables,
formatting worksheet output, generating APA write-ups, and building
interactive homework checkers.

## 1. Run the Analysis

`chi_square_answers()` performs a Pearson's chi-square test for two
categorical variables and returns observed/expected frequencies along with
the test statistic.

```{r}
#| label: run-chi
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# Run chi-square test
chi_result <- chi_square_answers(
  data = superman,
  var1 = "clark_grp",
  var2 = "tomatometer"
)

# Inspect the results
str(chi_result)
```

### What `chi_square_answers()` returns

| Element | Contents |
|---------|----------|
| `$ChiSquare` | `chi_sq`, `p_value`, `df`, `n`, `phi` |
| `$Observed` | Matrix of observed cell frequencies |
| `$Expected` | Matrix of expected cell frequencies |
| `$Effect_Size` | `phi` and `r` effect size |

## 2. Create Hypothesis Contingency Tables

Before running the analysis, students label a contingency table with
`<`, `>`, or `=` to predict the expected pattern. `create_rh_crosstabs()`
generates this table in both KEY and BLANK modes.

```{r}
#| label: rh-crosstabs
#| eval: false

# Blank version for students to fill in
create_rh_crosstabs(
  var1_name      = "Actor Height",
  var2_name      = "Tomatometer",
  var1_levels    = c("Under 6ft", "6ft or taller"),
  var2_levels    = c("Rotten", "Fresh"),
  KEY            = FALSE,
  include_totals = TRUE
)

# Answer key version with hypothesis pattern
create_rh_crosstabs(
  var1_name          = "Actor Height",
  var2_name          = "Tomatometer",
  var1_levels        = c("Under 6ft", "6ft or taller"),
  var2_levels        = c("Rotten", "Fresh"),
  KEY                = TRUE,
  chi_results        = chi_result,
  hypothesis_pattern = c("<", "="),
  include_totals     = TRUE
)
```

### Understanding the hypothesis pattern

The `hypothesis_pattern` is a character vector with one symbol per row of
the contingency table (one per level of `var1`). Each symbol describes
the expected relationship between the two columns of `var2`:

- `"<"` — the first column count is expected to be less than the second
- `">"` — the first column count is expected to be greater than the second
- `"="` — no expected difference between columns

## 3. Format Results for Worksheets

`format_chi2_results()` creates markdown output with the test statistic,
observed frequencies, and effect size. Toggle `KEY` for answer key vs.
blank student versions.

```{r}
#| label: format-chi
#| eval: false

# Answer key
chi_KEY <- format_chi2_results(
  rh_name          = "RH2",
  vars             = c("clark_grp", "tomatometer"),
  chi_results_list = chi_result,
  var1_labels      = c("Under 6ft", "6ft or taller"),
  var2_labels      = c("Rotten", "Fresh"),
  KEY              = TRUE
)

# Student worksheet (blanks)
chi_BLANK <- format_chi2_results(
  rh_name          = "RH2",
  vars             = c("clark_grp", "tomatometer"),
  chi_results_list = chi_result,
  var1_labels      = c("Under 6ft", "6ft or taller"),
  var2_labels      = c("Rotten", "Fresh"),
  KEY              = FALSE
)

# In your Quarto document (results: asis):
# cat(chi_KEY)
```

## 4. Create APA Contingency Tables

`create_apa_chi_crosstabs_table()` produces a publication-ready crosstab
as a `flextable`.

```{r}
#| label: apa-table
#| eval: false

# Answer key table
create_apa_chi_crosstabs_table(
  chi_results_list = chi_result,
  var1_name        = "Actor Height",
  var2_name        = "Tomatometer",
  var1_labels      = c("Under 6ft", "6ft or taller"),
  var2_labels      = c("Rotten", "Fresh"),
  KEY              = TRUE,
  table_title      = "Crosstabulation of Actor Height and Critic Ratings",
  table_number     = 2
)

# Blank version (pass NULL for chi_results_list)
create_apa_chi_crosstabs_table(
  chi_results_list = NULL,
  var1_name        = "Actor Height",
  var2_name        = "Tomatometer",
  var1_labels      = c("Under 6ft", "6ft or taller"),
  var2_labels      = c("Rotten", "Fresh"),
  KEY              = FALSE
)
```

## 5. Generate APA Write-Ups

`apa_chi_writeup()` generates a complete statistical write-up with
hypothesis evaluation and effect size interpretation.

```{r}
#| label: writeup
#| eval: false

writeup <- apa_chi_writeup(
  chi_results_list = chi_result,
  var1_name        = "actor height",
  var2_name        = "critic rating",
  var1_labels      = c("Under 6ft", "6ft or taller"),
  var2_labels      = c("Rotten", "Fresh"),
  hypothesis       = list(
    pattern              = c("=", "<"),
    rh_text              = "actors over 6ft tall would be more likely to appear in Fresh-rated media",
    comparison_var2_level = 2
  )
)

# cat(writeup)
```

### Hypothesis pattern for write-ups

The `hypothesis` list contains:

- `pattern` — character vector of `<`, `>`, `=` for each row of the table
- `rh_text` — plain-English description of the hypothesis
- `comparison_var2_level` — which column of `var2` is the "target" for
  the comparison (1 or 2)

## 6. Create Interactive Homework Checkers

`create_chisq_checker()` produces an HTML widget where students enter
observed frequencies and chi-square statistics.

```{r}
#| label: checker
#| eval: false

# In a Quarto HTML document:
create_chisq_checker(
  rh_name          = "RH2",
  chi_results_list = chi_result,
  var1_labels      = c("Under 6ft", "6ft or taller"),
  var2_labels      = c("Rotten", "Fresh")
)
```

## 7. Inline APA Statistics

```{r}
#| label: inline
#| eval: false

# Returns something like: χ²(1) = 4.52, *p* = .034
apa_inline_chi2(chi_result)
```

## 8. Effect Size Conversion

`pr_chi_to_r()` converts 2×2 cell frequencies to an r effect size,
useful for interpreting the practical significance of the association.

```{r}
#| label: effect-size
#| eval: false

# a, b, c, d are the four cells of the 2x2 table
pr_chi_to_r(a = 15, b = 10, c = 8, d = 17)
```

## Complete Lab Setup Example

```{r}
#| label: full-example
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# ── Analysis ─────────────────────────────────────────────
RH2_chi <- chi_square_answers(
  data = superman,
  var1 = "clark_grp",
  var2 = "tomatometer"
)

# ── Hypothesis Crosstabs ─────────────────────────────────
RH2_contingency_BLANK <- create_rh_crosstabs(
  var1_name = "Actor Height", var2_name = "Tomatometer",
  var1_levels = c("Under 6ft", "6ft or taller"),
  var2_levels = c("Rotten", "Fresh"),
  KEY = FALSE, include_totals = TRUE
)

RH2_contingency_KEY <- create_rh_crosstabs(
  var1_name = "Actor Height", var2_name = "Tomatometer",
  var1_levels = c("Under 6ft", "6ft or taller"),
  var2_levels = c("Rotten", "Fresh"),
  KEY = TRUE, chi_results = RH2_chi,
  hypothesis_pattern = c("<", "="),
  include_totals = TRUE
)

# ── Worksheet Output ─────────────────────────────────────
RH2_chi_BLANK <- format_chi2_results(
  rh_name = "RH2",
  vars = c("clark_grp", "tomatometer"),
  chi_results_list = RH2_chi,
  var1_labels = c("Under 6ft", "6ft or taller"),
  var2_labels = c("Rotten", "Fresh"),
  KEY = FALSE
)

RH2_chi_KEY <- format_chi2_results(
  rh_name = "RH2",
  vars = c("clark_grp", "tomatometer"),
  chi_results_list = RH2_chi,
  var1_labels = c("Under 6ft", "6ft or taller"),
  var2_labels = c("Rotten", "Fresh"),
  KEY = TRUE
)

# ── APA Tables ───────────────────────────────────────────
RH2_table_BLANK <- create_apa_chi_crosstabs_table(
  chi_results_list = NULL,
  var1_name = "Actor Height", var2_name = "Tomatometer",
  var1_labels = c("Under 6ft", "6ft or taller"),
  var2_labels = c("Rotten", "Fresh"),
  KEY = FALSE
)

RH2_table_KEY <- create_apa_chi_crosstabs_table(
  chi_results_list = RH2_chi,
  var1_name = "Actor Height", var2_name = "Tomatometer",
  var1_labels = c("Under 6ft", "6ft or taller"),
  var2_labels = c("Rotten", "Fresh"),
  KEY = TRUE,
  table_title = "Crosstabulation of Actor Height and Critic Ratings",
  table_number = 2
)

# ── Write-Up ─────────────────────────────────────────────
RH2_writeup <- apa_chi_writeup(
  chi_results_list = RH2_chi,
  var1_name = "actor height",
  var2_name = "critic rating",
  var1_labels = c("Under 6ft", "6ft or taller"),
  var2_labels = c("Rotten", "Fresh"),
  hypothesis = list(
    pattern = c("=", "<"),
    rh_text = "actors over 6ft would be more likely to appear in Fresh-rated media",
    comparison_var2_level = 2
  )
)
```

## Key Functions Reference

| Function | Purpose |
|----------|---------|
| `chi_square_answers()` | Run chi-square test, get χ², p, observed/expected |
| `create_rh_crosstabs()` | Hypothesis contingency table (KEY/BLANK) |
| `format_chi2_results()` | Markdown output for worksheets (KEY/BLANK) |
| `create_apa_chi_crosstabs_table()` | APA-style contingency table (flextable) |
| `apa_chi_writeup()` | Full APA write-up with hypothesis evaluation |
| `create_chisq_checker()` | Interactive HTML homework checker |
| `apa_inline_chi2()` | Inline APA-formatted χ² statistic |
| `pr_chi_to_r()` | Convert 2×2 cell frequencies to r effect size |
