---
title: "Correlations"
subtitle: "Pearson's r with descriptives, APA tables, and write-ups"
order: 2
---

## Overview

This guide shows how to use **psych350lab** to build a complete correlation
lab—from running the analysis through generating student worksheets, answer
keys, APA tables, write-ups, and interactive homework checkers.

Every code block below is a self-contained example you can copy into your
own lab setup file or Quarto document.

## 1. Run the Analysis

`corr_answers()` computes a Pearson correlation and returns descriptive
statistics for both variables. All numeric values are stored **unrounded**
so that display functions can format them appropriately.

```{r}
#| label: run-corr
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

# Load and prepare data
data(superman, package = "psych350data")

# Run the correlation
corr_result <- corr_answers(
  data = superman,
  var1 = "clark_height_in",
  var2 = "rt_audience_score"
)

# Inspect what comes back
str(corr_result)
```

### What `corr_answers()` returns

| Element | Contents |
|---------|----------|
| `$Correlation` | `r`, `p_value`, `df`, `n` |
| `$Descriptives` | Tibble with `variable`, `mean`, `sd`, `n` for each variable |

## 2. Format Results for Worksheets

`format_corr_results()` creates markdown-formatted output with two modes:
`Key = TRUE` fills in the answers; `Key = FALSE` leaves blanks for students.

Use these inside a Quarto chunk with `results: asis` so the markdown renders
properly.

```{r}
#| label: format-corr
#| eval: false

# Answer key version
corr_KEY <- format_corr_results(
  rh_name           = "RH1",
  vars              = c("clark_height_in", "rt_audience_score"),
  corr_results_list = corr_result,
  Key               = TRUE
)

# Student worksheet version (blanks)
corr_BLANK <- format_corr_results(
  rh_name           = "RH1",
  vars              = c("clark_height_in", "rt_audience_score"),
  corr_results_list = corr_result,
  Key               = FALSE
)

# In your Quarto document:
# cat(corr_KEY)   # for answer key
# cat(corr_BLANK) # for student worksheet
```

## 3. Create APA Correlation Tables

`create_corr_apa_table()` produces a publication-ready correlation matrix
as a `flextable`. The `KEY` argument toggles between filled and blank versions.

```{r}
#| label: apa-table
#| eval: false

# Answer key table
create_corr_apa_table(
  data       = superman,
  vars       = c("clark_height_in", "rt_audience_score"),
  var_labels = c("1. Actor Height (in)", "2. Audience Score"),
  KEY        = TRUE,
  table_title = "Table 1: Correlations Between Actor Height and Ratings"
)

# Blank table for worksheets (no data needed)
create_corr_apa_table(
  var_labels = c("1. Actor Height (in)", "2. Audience Score"),
  KEY        = FALSE
)
```

## 4. Generate APA Write-Ups

`apa_corr_writeup()` produces a complete statistical write-up with
hypothesis evaluation, descriptive statistics, and interpretation.

```{r}
#| label: writeup
#| eval: false

writeup <- apa_corr_writeup(
  corr_results_list    = corr_result,
  var1_label           = "actor height",
  var2_label           = "audience score",
  hypothesis           = list(
    direction = "positive",
    rh_text   = "taller actors would receive higher audience scores"
  ),
  include_descriptives = TRUE,
  table_number         = 1
)

# Use in a Quarto chunk with results: asis
# cat(writeup)
```

### Hypothesis directions

The `direction` field controls the write-up language:

- `"positive"` — expects r > 0
- `"negative"` — expects r < 0
- `"none"` — non-directional (two-tailed only)

## 5. Create Interactive Homework Checkers

`create_corr_checker()` produces an interactive HTML widget (using the
`webexercises` package) where students type in their answers and get
immediate feedback.

```{r}
#| label: checker
#| eval: false

# In a Quarto HTML document:
create_corr_checker(
  rh_name           = "RH1",
  vars              = c("clark_height_in", "rt_audience_score"),
  corr_results_list = corr_result
)
```

Students see fill-in-the-blank boxes for means, SDs, N, r, df, and p.
Correct answers turn green; incorrect answers turn red.

## 6. Inline APA Statistics

For quick inline reporting in Quarto text, use `apa_inline_r()`:

```{r}
#| label: inline
#| eval: false

# Returns something like: *r*(28) = .45, *p* = .012
apa_inline_r(corr_result)
```

## 7. SPSS-Style Scatterplots

Use `ggplot2` with `theme_SPSS()` and `number_SPSS` for plots that match
SPSS output:

```{r}
#| label: scatterplot
#| eval: false

library(ggplot2)

ggplot(superman, aes(x = clark_height_in, y = rt_audience_score)) +
  geom_point(
    shape = 21, fill = "#1192E8", color = "black",
    size = 2.5, stroke = 0.75, alpha = 0.7
  ) +
  theme_SPSS() +
  scale_x_continuous(labels = number_SPSS) +
  scale_y_continuous(labels = number_SPSS) +
  labs(x = "Actor Height (inches)", y = "Audience Score")
```

## Complete Lab Setup Example

Here is a minimal but complete setup file for a correlation lab section,
following the same pattern used in the actual course labs:

```{r}
#| label: full-example
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)
library(ggplot2)

# ── Data ─────────────────────────────────────────────────
data(superman, package = "psych350data")

# ── Analysis ─────────────────────────────────────────────
RH1_corr <- corr_answers(
  data = superman,
  var1 = "clark_height_in",
  var2 = "rt_audience_score"
)

RH1_vars <- c("clark_height_in", "rt_audience_score")

# ── Worksheet Output ─────────────────────────────────────
RH1_corr_BLANK <- format_corr_results(
  rh_name = "RH1", vars = RH1_vars,
  corr_results_list = RH1_corr, Key = FALSE
)

RH1_corr_KEY <- format_corr_results(
  rh_name = "RH1", vars = RH1_vars,
  corr_results_list = RH1_corr, Key = TRUE
)

# ── APA Tables ───────────────────────────────────────────
RH1_table_BLANK <- create_corr_apa_table(
  var_labels = c("1. Actor Height (in)", "2. Audience Score"),
  KEY = FALSE
)

RH1_table_KEY <- create_corr_apa_table(
  data = superman, vars = RH1_vars,
  var_labels = c("1. Actor Height (in)", "2. Audience Score"),
  KEY = TRUE,
  table_title = "Table 1: Correlations Between Actor Height and Ratings"
)

# ── Write-Up ─────────────────────────────────────────────
RH1_writeup <- apa_corr_writeup(
  corr_results_list = RH1_corr,
  var1_label = "actor height",
  var2_label = "audience score",
  hypothesis = list(
    direction = "positive",
    rh_text = "taller actors would receive higher audience scores"
  ),
  include_descriptives = TRUE
)

# ── Scatterplot ──────────────────────────────────────────
RH1_plot <- ggplot(superman, aes(x = clark_height_in, y = rt_audience_score)) +
  geom_point(shape = 21, fill = "#1192E8", color = "black",
             size = 2.5, stroke = 0.75, alpha = 0.7) +
  theme_SPSS() +
  scale_x_continuous(labels = number_SPSS) +
  scale_y_continuous(labels = number_SPSS) +
  labs(x = "Actor Height (in)", y = "Audience Score")
```

## Key Functions Reference

| Function | Purpose |
|----------|---------|
| `corr_answers()` | Run Pearson correlation, get r, p, df, descriptives |
| `format_corr_results()` | Markdown output for worksheets (KEY/BLANK) |
| `create_corr_apa_table()` | APA-style correlation matrix (flextable) |
| `apa_corr_writeup()` | Full APA write-up with hypothesis evaluation |
| `create_corr_checker()` | Interactive HTML homework checker |
| `apa_inline_r()` | Inline APA-formatted r statistic |
| `theme_SPSS()` / `number_SPSS` | SPSS-style plot theme and axis labels |
