---
title: "Factorial ANOVA (2×2)"
subtitle: "Between-groups, mixed, and within-groups 2×2 factorial designs"
order: 7
---

## Overview

**psych350lab** supports three types of 2×2 factorial ANOVA:

| Design | Function | When to use |
|--------|----------|-------------|
| **Between-groups (BG)** | `anova_factorial_answers()` | Both IVs are between-subjects |
| **Mixed (MG)** | `anova_factorial_mg_answers()` | One BG IV × one WG IV |
| **Within-groups (WG)** | `anova_factorial_wg_answers()` | Both IVs are within-subjects |

All three follow the same general workflow and return similarly structured
results. This guide covers each design with examples.

## Between-Groups Factorial (2×2 BG)

### 1. Run the Analysis

```{r}
#| label: run-bg-factorial
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# 2×2 BG factorial: Height group × Tomatometer → Audience score
bg_factorial <- anova_factorial_answers(
  data       = superman,
  dv         = "rt_audience_score",
  iv1        = "clark_grp",
  iv2        = "tomatometer",
  iv1_labels = c("Under 6ft", "6ft or taller"),
  iv2_labels = c("Rotten", "Fresh")
)

# Main effects and interaction
bg_factorial$ANOVA
# Cell means
bg_factorial$Descriptives
```

#### What `anova_factorial_answers()` returns

| Element | Contents |
|---------|----------|
| `$ANOVA` | List with `iv1`, `iv2`, and `interaction` sub-lists, each containing `F`, `p_value`, `df_effect`, `df_error`, `mse` |
| `$Descriptives` | Tibble with `iv1_level`, `iv2_level`, `mean`, `sd`, `n` for each cell |
| `$Marginal_Means` | Marginal means for each IV |

### 2. Create Checker Widgets

```{r}
#| label: bg-checkers
#| eval: false

# Factorial ANOVA statistics checker
create_factorial_anova_checker(
  rh_name            = "RH1",
  anova_results_list = bg_factorial,
  iv1_name           = "Height Group",
  iv2_name           = "Tomatometer"
)

# Cell descriptives checker
create_factorial_desc_checker(
  anova_results_list = bg_factorial,
  iv1_name           = "Height Group",
  iv2_name           = "Tomatometer"
)
```

### 3. Format Results

```{r}
#| label: bg-format
#| eval: false

# Complete formatted results
format_factorial_results(
  anova_results_list = bg_factorial,
  iv1_name           = "Height Group",
  iv2_name           = "Tomatometer",
  KEY                = TRUE
)
```

### 4. APA Tables

```{r}
#| label: bg-tables
#| eval: false

# Factorial ANOVA statistics table
create_factorial_anova_stats_table(
  rh_name            = "RH1",
  anova_results_list = bg_factorial,
  iv1_name           = "Height Group",
  iv2_name           = "Tomatometer"
)

# Cell means table
create_factorial_apa_desc_table(
  anova_results_list = bg_factorial,
  iv1_name           = "Height Group",
  iv2_name           = "Tomatometer",
  KEY                = TRUE,
  table_title        = "Cell Means for Audience Score"
)
```

### 5. Interaction and Main Effect Results

These functions generate formatted text describing the interaction and
main effects, suitable for worksheets:

```{r}
#| label: bg-effects
#| eval: false

# Interaction results
factorial_interaction_results(
  anova_results_list = bg_factorial,
  KEY                = TRUE
)

# Main effect results (specify which IV)
factorial_main_effect_results(
  anova_results_list = bg_factorial,
  iv_name            = "Height Group",
  which_iv           = "IV1",
  KEY                = TRUE
)
```

### 6. Write-Up

```{r}
#| label: bg-writeup
#| eval: false

writeup <- apa_2x2_bg_factorial_writeup(
  factorial_results_list = bg_factorial,
  dv_name                = "audience score",
  factor_a_name          = "height group",
  factor_b_name          = "Tomatometer rating",
  factor_a_labels        = c("under 6ft", "6ft or taller"),
  factor_b_labels        = c("Rotten", "Fresh"),
  hypothesis             = list(
    rh_text = "an interaction between height and rating on audience scores"
  ),
  report_main_effects    = TRUE
)

# cat(writeup)
```

## Mixed Factorial (2×2 MG)

### 1. Run the Analysis

In a mixed design, one IV is between-subjects and one is within-subjects
(repeated measures).

```{r}
#| label: run-mg-factorial
#| eval: false

# Mixed factorial: Height group (BG) × Score type (WG)
mg_factorial <- anova_factorial_mg_answers(
  data       = superman,
  dv         = "rt_audience_score",   # placeholder - see function docs
  iv1        = "clark_grp",           # between-subjects IV
  iv2        = "score_type",          # within-subjects IV
  iv1_labels = c("Under 6ft", "6ft or taller"),
  iv2_labels = c("Critics Score", "Audience Score")
)
```

### 2. Interaction and Main Effect Text

Mixed designs have their own formatting functions:

```{r}
#| label: mg-effects
#| eval: false

# Interaction results
factorial_mg_int_results(
  anova_results_list = mg_factorial,
  KEY                = TRUE
)

# Main effect results
factorial_mg_main_effect_results(
  anova_results_list = mg_factorial,
  iv_name            = "Height Group",
  which_iv           = "IV1",
  KEY                = TRUE
)
```

## Within-Groups Factorial (2×2 WG)

### 1. Run the Analysis

In a fully within-subjects design, each participant is measured in all
four cells. You provide the four DV column names directly.

```{r}
#| label: run-wg-factorial
#| eval: false

# Four columns: one per cell (IV1_level1 × IV2_level1, etc.)
wg_factorial <- anova_factorial_wg_answers(
  data       = superman,
  dv_a1b1    = "score_critics_fresh",
  dv_a1b2    = "score_critics_rotten",
  dv_a2b1    = "score_audience_fresh",
  dv_a2b2    = "score_audience_rotten",
  iv1_labels = c("Critics", "Audience"),
  iv2_labels = c("Fresh", "Rotten")
)
```

### 2. Interaction and Main Effect Text

```{r}
#| label: wg-effects
#| eval: false

# Interaction results
factorial_wg_int_results(
  anova_results_list = wg_factorial,
  KEY                = TRUE
)

# Main effect results
factorial_wg_main_effect_results(
  anova_results_list = wg_factorial,
  iv_name            = "Score Type",
  which_iv           = "IV1",
  KEY                = TRUE
)
```

### 3. Write-Up

```{r}
#| label: wg-writeup
#| eval: false

writeup <- apa_2x2_wg_factorial_writeup(
  wg_factorial_results_list = wg_factorial,
  dv_name                   = "rating score",
  factor_a_name             = "score type",
  factor_b_name             = "rating",
  factor_a_labels           = c("critics", "audience"),
  factor_b_labels           = c("Fresh", "Rotten"),
  hypothesis                = list(
    rh_text = "an interaction between score type and rating"
  )
)

# cat(writeup)
```

## Shared Functions Across All Factorial Designs

These work with all three factorial types:

```{r}
#| label: shared
#| eval: false

# Factorial ANOVA checker (interactive HTML)
create_factorial_anova_checker(
  rh_name = "RH1",
  anova_results_list = bg_factorial,
  iv1_name = "Height Group",
  iv2_name = "Tomatometer"
)

# Cell descriptives checker
create_factorial_desc_checker(
  anova_results_list = bg_factorial,
  iv1_name = "Height Group",
  iv2_name = "Tomatometer"
)

# Table with pairwise comparisons
factorial_table_with_comparisons(
  anova_results_list = bg_factorial,
  iv1_name = "Height Group",
  iv2_name = "Tomatometer"
)
```

## Key Functions Reference

### Analysis Functions

| Function | Design |
|----------|--------|
| `anova_factorial_answers()` | 2×2 between-groups |
| `anova_factorial_mg_answers()` | 2×2 mixed (1 BG × 1 WG) |
| `anova_factorial_wg_answers()` | 2×2 within-groups |

### Display Functions

| Function | Purpose |
|----------|---------|
| `create_factorial_anova_checker()` | Interactive statistics checker |
| `create_factorial_desc_checker()` | Interactive descriptives checker |
| `create_factorial_anova_stats_table()` | ANOVA statistics table |
| `create_factorial_apa_desc_table()` | Cell means table |
| `format_factorial_results()` | Formatted markdown output |
| `factorial_table_with_comparisons()` | Table with pairwise results |

### Effect Interpretation Functions

| Function | Design |
|----------|--------|
| `factorial_interaction_results()` | BG interaction text |
| `factorial_main_effect_results()` | BG main effect text |
| `factorial_mg_int_results()` | Mixed interaction text |
| `factorial_mg_main_effect_results()` | Mixed main effect text |
| `factorial_wg_int_results()` | WG interaction text |
| `factorial_wg_main_effect_results()` | WG main effect text |

### Write-Up Functions

| Function | Design |
|----------|--------|
| `apa_2x2_bg_factorial_writeup()` | BG factorial |
| `apa_2x2_wg_factorial_writeup()` | WG factorial |
