---
title: "Getting Started with psych350lab"
subtitle: "Installation, core concepts, and a quick tour"
order: 1
---

## What This Package Does

**psych350lab** helps instructors and TAs build statistics lab materials in
R that replicate SPSS/jamovi output. The package provides four types of
output for each statistical test:

1. **Worksheet output** — markdown-formatted results with a KEY/BLANK
   toggle for answer keys vs. student fill-in-the-blank worksheets
   (rendered via Quarto to `.docx`)
2. **APA tables** — publication-ready tables as `flextable` objects
3. **Statistical write-ups** — complete APA-style interpretation text
4. **Homework checkers** — interactive HTML widgets (via `webexercises`)
   where students enter answers and get immediate feedback

## Installation

Install both packages from GitHub using `pak`:

```{r}
#| label: install
#| eval: false

install.packages("pak")

pak::pak("emmarshall/psych350data")
pak::pak("emmarshall/psych350lab")
```

### GitHub Authentication (First Time Only)

```{r}
#| label: github-auth
#| eval: false

install.packages(c("gitcreds", "usethis"))

usethis::create_github_token()
gitcreds::gitcreds_set()
usethis::git_sitrep()
```

## Core Workflow

Every lab section follows the same five-step pattern, regardless of the
statistical test:

```{r}
#| label: workflow
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)
library(ggplot2)

# ── Step 1: Load and prepare data ────────────────────────
data(superman, package = "psych350data")

# ── Step 2: Run the analysis ─────────────────────────────
result <- corr_answers(
  data = superman,
  var1 = "clark_height_in",
  var2 = "rt_audience_score"
)

# ── Step 3: Format for worksheets (KEY or BLANK) ─────────
key_output <- format_corr_results(
  rh_name = "RH1",
  vars = c("clark_height_in", "rt_audience_score"),
  corr_results_list = result,
  Key = TRUE
)

# ── Step 4: Create APA table ─────────────────────────────
create_corr_apa_table(
  data = superman,
  vars = c("clark_height_in", "rt_audience_score"),
  var_labels = c("1. Actor Height (in)", "2. Audience Score"),
  KEY = TRUE
)

# ── Step 5: Generate write-up ────────────────────────────
writeup <- apa_corr_writeup(
  corr_results_list = result,
  var1_label = "actor height",
  var2_label = "audience score",
  hypothesis = list(
    direction = "positive",
    rh_text = "taller actors would receive higher audience scores"
  ),
  include_descriptives = TRUE
)
```

## The KEY/BLANK System

Most display functions have a `KEY` argument (sometimes called `Key`) that
toggles between two modes:

- **`KEY = TRUE`** — fills in all values (instructor answer key)
- **`KEY = FALSE`** — shows blanks for students to fill in

This lets you generate both versions from the same Quarto source file
using a `params` toggle:

```yaml
# In your .qmd YAML header:
params:
  hide_answers: true
```

```{r}
#| label: key-blank
#| eval: false

# In a code chunk:
format_corr_results(
  rh_name = "RH1",
  vars = c("clark_height_in", "rt_audience_score"),
  corr_results_list = result,
  Key = !params$hide_answers
)
```

## Available Analyses

The package follows the lab course progression. Each analysis type has its
own detailed guide:

| Lab Topic | Analysis | Guide |
|-----------|----------|-------|
| Bivariate | Pearson's correlation | [Correlations](correlations.qmd) |
| Bivariate | Chi-square (2×2) | [Chi-Square (2×2)](chi-square-2x2.qmd) |
| Group differences | One-way ANOVA (BG & WG) | [One-Way ANOVA](oneway-anova.qmd) |
| Group differences | K-group one-way ANOVA | [K-Group ANOVA](kgroup-anova.qmd) |
| Categorical | K-group chi-square (k×2) | [K-Group Chi-Square](kgroup-chi-square.qmd) |
| Factorial | 2×2 factorial ANOVA (BG, MG, WG) | [Factorial ANOVA](factorial-anova.qmd) |
| Multivariate | Multiple linear regression | [Regression](regression.qmd) |

## Function Naming Conventions

Functions follow a consistent naming pattern:

| Pattern | Purpose | Example |
|---------|---------|---------|
| `*_answers()` | Run analysis, return raw results | `corr_answers()`, `bg_anova_answers()` |
| `format_*_results()` | Markdown for worksheets (KEY/BLANK) | `format_corr_results()` |
| `create_*_table()` | APA flextable | `create_corr_apa_table()` |
| `create_*_checker()` | Interactive HTML checker | `create_corr_checker()` |
| `apa_*_writeup()` | Full APA write-up text | `apa_corr_writeup()` |
| `apa_inline_*()` | Inline APA statistic | `apa_inline_r()` |
| `format_*()` | Format a single value | `format_p_value()`, `format_stat()` |

## Formatting Utilities

These helper functions format individual values for display:

```{r}
#| label: format-helpers
#| eval: false

format_p_value(0.023)                         # ".023"
format_p_value(0.0004)                        # "< .001"
format_p_value(0.023, include_p = TRUE)       # "p = .023"
format_stat(3.456)                            # "3.46"
format_stat(0.456, remove_leading_zero = TRUE) # ".46"
format_r(0.456)                               # ".46"
format_effect(0.35)                           # ".35"
format_int(25.4)                              # "25"
p_to_stars(0.003)                             # "**"
key_or_blank(0.45, KEY = TRUE, format_fn = format_r) # ".45"
key_or_blank(0.45, KEY = FALSE)               # "______"
```

## SPSS-Style Plots

`theme_SPSS()` and `number_SPSS` replicate the SPSS chart appearance:

```{r}
#| label: plot-theme
#| eval: false

ggplot(superman, aes(x = clark_height_in, y = rt_audience_score)) +
  geom_point(
    shape = 21, fill = "#1192E8", color = "black",
    size = 2.5, stroke = 0.75, alpha = 0.7
  ) +
  theme_SPSS() +
  scale_x_continuous(labels = number_SPSS) +
  scale_y_continuous(labels = number_SPSS) +
  labs(x = "Actor Height (inches)", y = "Audience Score")
```

## Typical Lab File Structure

A complete lab consists of three files:

```
labs/lab-07/
├── lab07-setup.R          # Data prep + all analyses
├── lab07_worksheet.qmd    # Quarto → Word worksheet
└── lab07_checker.qmd      # Quarto → HTML homework checker
```

The **setup file** runs all analyses and stores results in objects.
The **worksheet** and **checker** files source the setup file and display
the results using the appropriate output functions.

See the [Correlations](correlations.qmd) guide for a complete setup file
example.
