---
title: "Homework Checkers"
subtitle: "Interactive webexercises answer checkers for every analysis type"
order: 10
format:
  html:
    mainfont: Sora
    filters:
      - _extensions/nebraska/hwchecker/webexercises.lua
    theme:
      - default
      - _extensions/nebraska/hwchecker/_variables.scss
      - _extensions/nebraska/hwchecker/_mixins.scss
      - _extensions/nebraska/hwchecker/custom.scss
    css:
      - _extensions/nebraska/hwchecker/webex.css
    include-after-body:
      - _extensions/nebraska/hwchecker/webex.js
execute: 
  warning: false
  message: false
---

## Overview

The `psych350lab` package provides interactive homework checkers that
students can use in Quarto HTML documents. Students enter their answers
into fill-in-the-blank boxes or select from dropdown menus, and the
boxes turn **green** for correct or **red** for incorrect.

These checkers rely on the `webexercises` and `tinytable` packages.

## Setup

This vignette uses the homework-checker theme and webexercises support
bundled inside the `nebraska` Quarto extension
(`_extensions/nebraska/hwchecker/`). To use it in your own documents,
make sure the `_extensions/nebraska` folder is in your project and add
the hwchecker assets to your YAML header:

```yaml
format:
  html:
    mainfont: Sora
    filters:
      - _extensions/nebraska/hwchecker/webexercises.lua
    theme:
      - default
      - _extensions/nebraska/hwchecker/_variables.scss
      - _extensions/nebraska/hwchecker/_mixins.scss
      - _extensions/nebraska/hwchecker/custom.scss
    css:
      - _extensions/nebraska/hwchecker/webex.css
    include-after-body:
      - _extensions/nebraska/hwchecker/webex.js
```

Every checker chunk must include `results: asis` so that the HTML renders
correctly:

```yaml
#| results: asis
```

## General Workflow

All checkers follow the same two-step pattern:

1. **Compute results** using an `*_answers()` function
2. **Create checker** using a `create_*_checker()` function

```r
# Step 1: Compute
result <- some_answers(data, ...)

# Step 2: Create checker
create_some_checker(result, ...)
```

### How Rounding and Computation Work

The `*_answers()` functions return **unrounded** values from the
underlying statistical tests. No rounding is applied at the computation
step so that full precision is preserved. The `create_*_checker()`
functions then round each statistic to the appropriate number of decimal
places (e.g., 2 for means and standard deviations, 3 for correlations
and *p*-values) before building the fill-in-the-blank widgets. This
means the checker's expected answers always match what students would
report following APA formatting conventions.

All analyses follow SPSS defaults. Missing values coded as −99 (or other
SPSS user-missing codes) are converted to `NA` when the data are read
with `haven::read_sav()` and `haven::zap_missing()`, so they are
automatically excluded from computations.

## Shared Setup

All examples below use the Superman media dataset. Load the data once at
the top of your document:

```{r}
#| label: setup
#| message: false
library(psych350lab)
library(dplyr)
data(superman, package = "psych350data")

# Derive decade from year (not in the base dataset)
superman <- superman |>
  mutate(
    decade = case_when(
      year < 1970 ~ 1,
      year < 2000 ~ 2,
      year < 2010 ~ 3,
      TRUE        ~ 4
    )
  )
```


## 1. Descriptive Statistics

Students enter Mean, SD, and SEM for each variable and identify whether
the mean is interpretable.

### Compute

```{r}
#| label: desc-compute

# Select and filter your analysis variables
my_data <- superman |>
  select(num, year, clark_height_in, clark_grp, height_diff, height_gap) |>
  filter(!is.na(height_diff))

# Compute descriptive statistics
desc_result <- descriptives_answers(my_data,
  vars = c("num", "year", "clark_height_in", "clark_grp",
           "height_diff", "height_gap")
)
```

### Checker

```{r}
#| label: desc-checker
#| results: asis

create_descriptives_checker(
  vars = c("num", "year", "clark_height_in", "clark_grp",
           "height_diff", "height_gap"),
  desc_results_list = desc_result,
  label          = "num",
  quantitative   = c("year", "clark_height_in", "height_diff"),
  binary         = "clark_grp",
  multi_category = "height_gap",
  var_labels = c(
    num            = "Film Number",
    year           = "Release Year",
    clark_height_in = "Clark Height (in)",
    clark_grp      = "Height Group",
    height_diff    = "Height Difference",
    height_gap     = "Height Gap Category"
  )
)
```

### Variable Type Categories

| Category         | Meaning                          | Mean Interpretable? |
|------------------|----------------------------------|---------------------|
| `label`          | ID variable (e.g., film number)  | No                  |
| `quantitative`   | Continuous numeric variable      | Yes                 |
| `binary`         | Dichotomous variable (2 levels)  | Yes                 |
| `multi_category` | Nominal variable (3+ levels)     | No                  |


## 2. Correlations

Students report the Pearson *r*, *p*-value, degrees of freedom, and
descriptive statistics for a bivariate correlation.

### Compute

```{r}
#| label: corr-compute

corr_result <- corr_answers(superman, "clark_height_in", "rt_critics_score")
```

### Checker

```{r}
#| label: corr-checker
#| results: asis

create_corr_checker(
  rh_name = "RH1",
  vars = c("Clark Height", "Critics Score"),
  corr_results_list = corr_result
)
```


## 3. Chi-Square (2×2)

Students report observed frequencies, the chi-square statistic,
*p*-value, and degrees of freedom for a 2×2 test of independence.

### Compute

```{r}
#| label: chi2x2-compute

chi_result <- chi_square_answers(superman, "clark_grp", "tomatometer")
```

### Checker

```{r}
#| label: chi2x2-checker
#| results: asis

create_chisq_checker(
  rh_name      = "RH1",
  chi_results_list = chi_result,
  var1_labels  = c("Under 6ft", "6ft+"),
  var2_labels  = c("Rotten", "Fresh")
)
```


## 4. K-Group Chi-Square (k×2)

When the row variable has three or more levels, a significant omnibus
chi-square is followed by pairwise 2×2 comparisons. Two checkers are
needed: one for the omnibus test and one for the pairwise follow-ups.

### Compute

```{r}
#| label: kgroup-chi-compute

kgroup_chi <- chi_square_kgroup_answers(
  data       = superman,
  var1       = "decade",
  var2       = "tomatometer",
  var1_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels = c("Rotten", "Fresh")
)
```

### Omnibus Checker

```{r}
#| label: kgroup-chi-omnibus-checker
#| results: asis

create_chisq_omnibus_table(
  rh_name            = "RH1",
  chisq_results_list = kgroup_chi,
  var1_labels        = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels        = c("Rotten", "Fresh")
)
```

### Pairwise Checker

```{r}
#| label: kgroup-chi-pairwise-checker
#| results: asis

create_chisq_pairwise_checker(
  chisq_results_list = kgroup_chi,
  var1_labels        = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels        = c("Rotten", "Fresh")
)
```


## 5. One-Way Between-Groups ANOVA (2 groups)

Students report the *F* statistic, *p*-value, degrees of freedom, MSE,
and group descriptives for a two-group between-groups comparison.

### Compute

```{r}
#| label: bg-anova-compute

bg_result <- bg_anova_answers(superman, iv = "clark_grp", dv = "rt_critics_score")
```

### Checker

```{r}
#| label: bg-anova-checker
#| results: asis

create_bg_anova_checker(
  rh_name            = "RH1",
  vars               = c("clark_grp", "rt_critics_score"),
  anova_results_list = bg_result
)
```


## 6. One-Way Within-Groups ANOVA (2 conditions)

Students report the *F* statistic, *p*-value, degrees of freedom, MSE,
and condition descriptives for a repeated-measures comparison.

### Compute

```{r}
#| label: wg-anova-compute

wg_result <- wg_anova_answers(
  superman,
  dv1 = "rt_critics_score",
  dv2 = "rt_audience_score",
  dv1_label = "Critics Score",
  dv2_label = "Audience Score"
)
```

### Checker

```{r}
#| label: wg-anova-checker
#| results: asis

create_wg_anova_checker(
  rh_name            = "RH1",
  vars               = c("Critics Score", "Audience Score"),
  anova_results_list = wg_result,
  condition_labels   = c("Critics Score", "Audience Score")
)
```


## 7. K-Group Between-Groups ANOVA (3+ groups)

When the IV has three or more groups, the omnibus *F*-test is followed
by LSD pairwise comparisons. Two checkers are needed.

### Compute

```{r}
#| label: kgroup-bg-compute

kgroup_bg <- anova_kgroup_answers(
  data         = superman,
  dv           = "rt_critics_score",
  iv           = "decade",
  group_labels = c("1950s", "1970-80s", "2000s", "2010s")
)
```

### Omnibus Checker

```{r}
#| label: kgroup-bg-omnibus-checker
#| results: asis

create_anova_omnibus_checker(
  rh_name            = "RH1",
  anova_results_list = kgroup_bg,
  group_labels       = c("1950s", "1970-80s", "2000s", "2010s")
)
```

### LSD Pairwise Checker

```{r}
#| label: kgroup-bg-pairwise-checker
#| results: asis

create_lsd_pairwise_checker(
  anova_results_list = kgroup_bg,
  group_labels       = c("1950s", "1970-80s", "2000s", "2010s")
)
```


## 8. Factorial ANOVA (2×2 Between-Groups / Mixed)

Students report the interaction and main-effect *F*-tests, degrees of
freedom, MSE, and follow-up comparisons. Two checkers cover the ANOVA
results and the cell/marginal descriptive statistics.

### Compute

```{r}
#| label: factorial-compute

sm_factorial <- superman |>
  mutate(era = if_else(year >= 2000, "Post-2000", "Pre-2000"))

factorial_result <- anova_factorial_answers(
  data      = sm_factorial,
  dv        = "rt_critics_score",
  iv1       = "clark_grp",
  iv2       = "era",
  iv1_labels = c("Under 6ft", "6ft+"),
  iv2_labels = c("Pre-2000", "Post-2000")
)
```

### ANOVA Checker

```{r}
#| label: factorial-anova-checker
#| results: asis

create_factbg_anova_checker(
  rh_name            = "RH1",
  anova_results_list = factorial_result,
  iv1_name           = "Height Group",
  iv2_name           = "Era"
)
```

### Descriptives Checker (cell means and EMMs)

```{r}
#| label: factorial-desc-checker
#| results: asis

create_factbg_desc_checker(
  anova_results_list = factorial_result,
  iv1_name           = "Height Group",
  iv2_name           = "Era",
  iv1_labels         = c("Under 6ft", "6ft+"),
  iv2_labels         = c("Pre-2000", "Post-2000")
)
```


## 9. Linear Regression

Students report model-level statistics (*R*, *R*², *F*, df, *p*) and
predictor-level statistics (*r*, *b*, significance, interpretation
category). Two checkers split these.

### Compute

```{r}
#| label: reg-compute

sm_reg <- superman |>
  filter(
    !is.na(rt_critics_score),
    !is.na(clark_height_in),
    !is.na(rt_audience_score)
  )

reg_result <- linear_reg_answers(
  data              = sm_reg,
  criterion         = "rt_critics_score",
  quant_predictors  = c("clark_height_in", "rt_audience_score"),
  quant_labels      = c("Clark Height", "Audience Score"),
  criterion_label   = "Critics Score"
)
```

### Model Summary Checker

```{r}
#| label: reg-model-checker
#| results: asis

create_regression_model_checker(reg_result)
```

### Predictor Results Checker

```{r}
#| label: reg-predictor-checker
#| results: asis

create_regression_predictor_checker(reg_result, show_legend = TRUE)
```


## Quick Reference

### Analysis Functions

| Function                         | Analysis Type        | Key Arguments                                       |
|----------------------------------|----------------------|-----------------------------------------------------|
| `descriptives_answers()`         | Univariate stats     | `data`, `vars`                                      |
| `corr_answers()`                 | Correlation          | `data`, `var1`, `var2`                              |
| `chi_square_answers()`           | Chi-Square (2×2)     | `data`, `var1`, `var2`                              |
| `chi_square_kgroup_answers()`    | Chi-Square (k×2)     | `data`, `var1`, `var2`, `var1_labels`, `var2_labels` |
| `bg_anova_answers()`             | BG ANOVA (2 groups)  | `data`, `iv`, `dv`                                  |
| `wg_anova_answers()`             | WG ANOVA (2 cond.)   | `data`, `dv1`, `dv2`                                |
| `anova_kgroup_answers()`         | BG ANOVA (k groups)  | `data`, `dv`, `iv`, `group_labels`                  |
| `anova_factorial_answers()`      | 2×2 Factorial (BG)   | `data`, `dv`, `iv1`, `iv2`                          |
| `anova_factmg_answers()`         | 2×2 Factorial (MG)   | `data`, `dv`, `iv1`, `iv2`                          |
| `anova_factwg_answers()`         | 2×2 Factorial (WG)   | `data`, `dv_a1b1`, `dv_a1b2`, `dv_a2b1`, `dv_a2b2` |
| `linear_reg_answers()`           | Regression           | `data`, `criterion`, `quant_predictors`             |

### Checker Functions

| Function                              | Analysis Type        | Key Components                          |
|---------------------------------------|----------------------|-----------------------------------------|
| `create_descriptives_checker()`       | Univariate           | Mean, SD, SEM, interpretability         |
| `create_corr_checker()`               | Correlation          | *r*, *p*, df, descriptives              |
| `create_chisq_checker()`              | Chi-Square (2×2)     | χ², *p*, df, frequencies                |
| `create_chisq_omnibus_table()`        | Chi-Square omnibus   | χ², *p*, df, *N*, frequencies           |
| `create_chisq_pairwise_checker()`     | Chi-Square pairwise  | χ² critical, percentages, effect sizes  |
| `create_bg_anova_checker()`           | BG ANOVA (2 groups)  | *F*, *p*, df, MSE, group stats          |
| `create_wg_anova_checker()`           | WG ANOVA (2 cond.)   | *F*, *p*, df, MSE, condition stats      |
| `create_anova_omnibus_checker()`      | BG ANOVA (k groups)  | *F*, *p*, df, MSE, *N*, *k*, group stats |
| `create_lsd_pairwise_checker()`       | LSD pairwise         | LSD MMD, mean differences, effect sizes |
| `create_factbg_anova_checker()`       | 2×2 Factorial (BG) ANOVA | Interaction, main effects, LSD          |
| `create_factbg_desc_checker()`        | 2×2 Factorial (BG) desc. | Cell means, EMMs                        |
| `create_factmg_anova_checker()`       | 2×2 Factorial (MG) ANOVA | Interaction, main effects, LSD          |
| `create_factmg_desc_checker()`        | 2×2 Factorial (MG) desc. | Cell means, EMMs                        |
| `create_factwg_anova_checker()`       | 2×2 Factorial (WG) ANOVA | Interaction, main effects, LSD          |
| `create_factwg_desc_checker()`        | 2×2 Factorial (WG) desc. | Cell means, EMMs                        |
| `create_regression_model_checker()`   | Regression model     | *R*, *R*², *F*, df, *p*                 |
| `create_regression_predictor_checker()` | Regression predictors | *r*, *b*, significance, categories    |


## Troubleshooting

**Tables not rendering.** Make sure your chunk includes `#| results: asis`.

**webexercises styles missing.** Check that your YAML includes the CSS
and JS files as shown in the Setup section above.

**Checker shows wrong answers.** Make sure you are passing the result
from the `*_answers()` function, not raw data.

**Package not found errors.** Install required packages with
`install.packages(c("tinytable", "webexercises"))`.
