---
title: "Creating Homework Checkers with psych350lab"
format:
  hwchecker-html: default
execute: 
  echo: false
---

## Overview

The `psych350lab` package provides interactive homework checkers that students can use in Quarto HTML documents. Students enter their answers into fill-in-the-blank boxes (`fitb()`) or select from dropdown menus (`mcq()`), and the boxes turn **green** for correct or **red** for incorrect.

These checkers rely on the `webexercises` and `tinytable` packages.

## Setting Up a Quarto Document

Your Quarto HTML document needs the webexercises CSS and JS files. Add this to your YAML header:
```yaml
format:
  html:
    css: [webexercises.css]
    include-after-body: [webexercises.js]
```

You can get these files by running `webexercises::create_quarto_doc()` or copying them from the webexercises package installation.

## General Workflow

All checkers follow the same two-step pattern:

1. **Compute results** using a `*_answers()` function
2. **Create checker** using a `create_*_checker()` function
```r
# Step 1: Compute
result <- some_answers(data, ...)

# Step 2: Create checker
create_some_checker(result, ...)
```

## Lab 06: Descriptive Statistics Checker

This is the most common checker — students enter Mean, SD, and SEM for each variable and identify whether the mean is interpretable.

### Step 1: Setup and prepare data
```{r}
#| label: setup
#| message: false
library(psych350lab)
library(dplyr)
data(superman)

my_data <- superman |> 
  select(num, year, clark_height_in, clark_grp, height_diff, height_gap) |>
  filter(!is.na(height_diff))
```

### Step 2: Compute descriptive statistics
```{r}
#| label: descriptives-compute
desc_result <- descriptives_answers(my_data,
  vars = c("num", "year", "clark_height_in", "clark_grp", "height_diff", "height_gap")
)

# View the results
desc_result$Descriptives
desc_result$Sample_Size
```

### Step 3: Create the checker
```{r}
#| label: descriptives-checker
#| results: asis

create_descriptives_checker(
  vars = c("num", "year", "clark_height_in", "clark_grp", "height_diff", "height_gap"),
  desc_results_list = desc_result,
  label = "num",
  quantitative = c("year", "clark_height_in", "height_diff"),
  binary = "clark_grp",
  multi_category = "height_gap",
  var_labels = c(
    num = "Film Number",
    year = "Release Year", 
    clark_height_in = "Clark Height (in)",
    clark_grp = "Height Group",
    height_diff = "Height Difference",
    height_gap = "Height Gap Category"
  )
)
```

### Variable Type Categories

| Category | Meaning | Mean Interpretable? |
|----------|---------|---------------------|
| `label` | ID variable (e.g., participant number) | No |
| `quantitative` | Continuous numeric variable | Yes |
| `binary` | Dichotomous variable (2 levels) | Yes |
| `multi_category` | Nominal variable (3+ levels) | No |


## Lab 07: Correlation Checker

Create an interactive checker for Pearson correlation results:
```{r}
#| label: correlation-checker
#| results: asis

# Step 1: Compute correlation
corr_result <- corr_answers(superman, "clark_height_in", "rt_critics_score")

# Step 2: Create checker
create_corr_checker(
  rh_name = "RH1",
  vars = c("Clark Height", "Critics Score"),
  corr_results_list = corr_result
)
```


## Lab 07: Chi-Square Checker

For categorical data analysis:
```{r}
#| label: chisq-checker
#| results: asis

# Step 1: Compute chi-square
chi_result <- chi_square_answers(superman, "clark_grp", "tomatometer")

# Step 2: Create checker
create_chisq_checker(
  rh_name = "RH1",
  chi_results_list = chi_result,
  var1_labels = c("Under 6ft", "6ft+"),
  var2_labels = c("Rotten", "Fresh")
)
```


## Lab 08: Between-Groups ANOVA Checker

For comparing two groups:
```{r}
#| label: bg-anova-checker
#| results: asis

# Step 1: Compute ANOVA
anova_result <- bg_anova_answers(superman, iv = "clark_grp", dv = "rt_critics_score")

# Step 2: Create checker
create_bg_anova_checker(
  rh_name = "RH1",
  vars = c("clark_grp", "rt_critics_score"),
  anova_results_list = anova_result
)
```


## Lab 08: Within-Groups ANOVA Checker

For repeated measures designs:
```{r}
#| label: wg-anova-checker
#| results: asis

# Step 1: Compute repeated measures ANOVA
wg_result <- wg_anova_answers(
  superman,
  dv1 = "rt_critics_score",
  dv2 = "rt_audience_score"
)

# Step 2: Create checker
create_wg_anova_checker(
  rh_name = "RH1",
  vars = c("rt_critics_score", "rt_audience_score"),
  anova_results_list = wg_result
)
```


## Lab 11: Multi-Group ANOVA with Pairwise Comparisons

For designs with 3+ groups:
```{r}
#| label: multigroup-setup
#| eval: false

# Step 1: Compute multi-group ANOVA
multigroup_result <- anova_kgroup_answers(
  superman,
  dv = "rt_critics_score",
  iv = "type",
  group_labels = c("Film", "TV", "Serial")
)
```

### Omnibus F-Test Checker
```{r}
#| label: omnibus-checker
#| results: asis
#| eval: false

# Step 2a: Create omnibus checker
create_anova_omnibus_checker("RH1", multigroup_result)
```

### LSD Pairwise Comparisons Checker
```{r}
#| label: lsd-checker
#| results: asis
#| eval: false

# Step 2b: Create pairwise checker
create_lsd_pairwise_checker(multigroup_result)
```


## Lab 12: Factorial ANOVA Checker

For 2×2 factorial designs:
```{r}
#| label: factorial-setup

# Prepare data with second grouping variable
sm_factorial <- superman |>
  mutate(era = if_else(year >= 2000, "Post-2000", "Pre-2000"))
```
```{r}
#| label: factorial-checker
#| results: asis
#| eval: false

# Step 1: Compute factorial ANOVA
factorial_result <- anova_factorial_answers(
  sm_factorial,
  dv = "rt_critics_score",
  iv1 = "clark_grp",
  iv2 = "era",
  iv1_labels = c("Under 6ft", "6ft+"),
  iv2_labels = c("Pre-2000", "Post-2000")
)

# Step 2: Create checker
create_factorial_anova_checker(
  rh_name = "RH1",
  anova_results_list = factorial_result,
  iv1_name = "Height Group",
  iv2_name = "Era"
)
```


## Lab 13: Regression Checkers

### Setup and compute regression
```{r}
#| label: regression-setup

# Prepare data
sm <- superman |>
  filter(
    !is.na(rt_critics_score),
    !is.na(clark_height_in),
    !is.na(rt_audience_score)
  )

# Step 1: Compute regression
reg_result <- linear_reg_answers(
  data = sm,
  criterion = "rt_critics_score",
  quant_predictors = c("clark_height_in", "rt_audience_score"),
  quant_labels = c("Clark Height", "Audience Score"),
  criterion_label = "Critics Score"
)
```

### Model Summary Checker
```{r}
#| label: regression-model-checker
#| results: asis

# Step 2a: Create model checker
create_regression_model_checker(reg_result)
```

### Predictor Results Checker
```{r}
#| label: regression-predictor-checker
#| results: asis

# Step 2b: Create predictor checker
create_regression_predictor_checker(reg_result, show_legend = TRUE)
```


## Quick Reference

### Analysis Functions (`*_answers()`)

| Function | Analysis Type | Key Arguments |
|----------|---------------|---------------|
| `descriptives_answers()` | Univariate stats | `data`, `vars` |
| `corr_answers()` | Correlation | `data`, `var1`, `var2` |
| `chi_square_answers()` | Chi-Square | `data`, `var1`, `var2` |
| `bg_anova_answers()` | BG ANOVA (2 groups) | `data`, `iv`, `dv` |
| `wg_anova_answers()` | WG ANOVA (2 conditions) | `data`, `dv1`, `dv2` |
| `anova_kgroup_answers()` | ANOVA (k groups) | `data`, `dv`, `iv`, `group_labels` |
| `anova_factorial_answers()` | 2×2 Factorial | `data`, `dv`, `iv1`, `iv2` |
| `linear_reg_answers()` | Regression | `data`, `criterion`, `quant_predictors` |

### Checker Functions (`create_*_checker()`)

| Function | Analysis Type | Key Components |
|----------|---------------|----------------|
| `create_descriptives_checker()` | Univariate | Mean, SD, SEM, interpretability |
| `create_corr_checker()` | Correlation | r, p, df, descriptives |
| `create_chisq_checker()` | Chi-Square (2×2) | χ², p, df, frequencies |
| `create_chisq_omnibus_checker()` | Chi-Square (k groups) | χ², p, df, N, frequencies |
| `create_chisq_pairwise_checker()` | Chi-Square pairwise | χ² critical, percentages, effect sizes |
| `create_bg_anova_checker()` | BG ANOVA (2 groups) | F, p, df, MSE, group stats |
| `create_wg_anova_checker()` | WG ANOVA (2 conditions) | F, p, df, MSE, condition stats |
| `create_anova_omnibus_checker()` | ANOVA (k groups) | F, p, df, MSE, N, k, group stats |
| `create_lsd_pairwise_checker()` | LSD pairwise | LSD MMD, mean differences, effect sizes |
| `create_factorial_anova_checker()` | 2×2 Factorial | Interaction, main effects, LSD |
| `create_factorial_desc_checker()` | 2×2 Factorial | Cell means, EMMs |
| `create_regression_model_checker()` | Regression | R, R², F, df, p |
| `create_regression_predictor_checker()` | Regression | r, b, significance, categories |


## Styling Tips

### Custom Table Width

Adjust the table width using tinytable styling:
```{r}
#| label: styling-example
#| eval: false

create_corr_checker("RH1", vars, result) |>
  tinytable::style_tt(
    bootstrap_css_rule = "width: 90%; margin: auto;"
  )
```

### Combining Multiple Checkers

Stack multiple checkers in sequence:
```{r}
#| label: combined-checkers
#| results: asis
#| eval: false

# Correlation section
cat("### Correlation Results\n\n")
create_corr_checker("RH1", vars1, corr_result)

cat("\n\n### ANOVA Results\n\n")
create_bg_anova_checker("RH2", vars2, anova_result, group_labels)
```


## Troubleshooting

### Tables Not Rendering

Ensure you have `results: asis` in your chunk options:
```yaml
#| results: asis
```

### webexercises Styles Missing

Check that your YAML includes the CSS and JS files:
```yaml
format:
  html:
    css: [webexercises.css]
    include-after-body: [webexercises.js]
```

### Package Not Found Errors

Install required packages:
```{r}
#| eval: false

install.packages(c("tinytable", "webexercises"))
```

### Checker Shows Wrong Answers

Make sure you're passing the result from the `*_answers()` function, not raw data:
```r
# Wrong
create_corr_checker("RH1", vars, superman)

# Correct
result <- corr_answers(superman, "var1", "var2")
create_corr_checker("RH1", vars, result)
```
