---
title: "K-Group One-Way ANOVA"
subtitle: "Between-groups ANOVA with 3+ groups, LSD/HSD post-hoc comparisons"
order: 5
---

## Overview

When the independent variable has three or more groups, a significant
omnibus F-test must be followed by pairwise comparisons. **psych350lab**
handles the full workflow: omnibus ANOVA, LSD and Tukey's HSD minimum
mean differences, pairwise comparisons with effect sizes, and power
assessment.

## 1. Run the Analysis

`anova_kgroup_answers()` performs the omnibus ANOVA, calculates
LSD/HSD thresholds, and runs all pairwise comparisons automatically.

```{r}
#| label: run-kgroup
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# Run k-group ANOVA
kgroup_result <- anova_kgroup_answers(
  data         = superman,
  dv           = "box_office_mil",
  iv           = "decade",
  group_labels = c("1950s", "1970-80s", "2000s", "2010s")
)

# Omnibus results
kgroup_result$ANOVA

# Group descriptives
kgroup_result$Descriptives

# LSD/HSD thresholds
kgroup_result$LSD

# Pairwise comparisons
kgroup_result$Pairwise
```

### What `anova_kgroup_answers()` returns

| Element | Contents |
|---------|----------|
| `$ANOVA` | `F`, `p_value`, `df_between`, `df_within`, `mse`, `total_n`, `k`, `mean_n` |
| `$Descriptives` | Tibble with `iv`, `mean`, `sd`, `n`, `group_label` per group |
| `$LSD` | `lsd_mmd`, `hsd_mmd`, `need_posthoc` (TRUE if p < .05) |
| `$Pairwise` | List of comparison results, each with `comparison`, `mean_diff`, `lsd_result`, `error_type`, `effect_size`, `power_problem` |
| `$group_labels` | Character vector of display labels |

## 2. Understanding Pairwise Comparisons

Each element in `$Pairwise` contains:

- **`comparison`** — which groups are being compared (e.g., "1950s vs 2000s")
- **`mean_diff`** — raw difference between group means
- **`lsd_result`** — `">"`, `"<"`, or `"="` based on whether |mean_diff| > LSD MMD
- **`error_type`** — "Type I & III" if significant, "Type II" if not
- **`effect_size`** — r effect size for the comparison
- **`power_problem`** — assessment of whether non-significant results may
  reflect a power issue

## 3. LSD and HSD Calculation

`lsd_hsd_calculator()` is called internally but can also be used
directly:

```{r}
#| label: lsd-hsd
#| eval: false

lsd_hsd_calculator(
  k              = 4,        # number of groups
  n_per_group    = 8,        # average n per group
  mse            = 1250.5,   # mean square error from omnibus ANOVA
  df_error_input = 28        # df_within from omnibus ANOVA
)
```

## 4. Format Results for Worksheets

The k-group ANOVA uses the same `format_bg_anova_results()` function
for the omnibus portion, plus additional pairwise display functions.

```{r}
#| label: format-kgroup
#| eval: false

# Omnibus results for worksheet
format_bg_anova_results(
  rh_name            = "RH1",
  vars               = c("Decade", "Box Office (millions)"),
  anova_results_list = kgroup_result,
  iv_labels          = c("1950s", "1970-80s", "2000s", "2010s"),
  KEY                = TRUE
)
```

## 5. Create APA Tables

```{r}
#| label: apa-tables
#| eval: false

# Descriptives table with group means
create_bg_anova_table(
  anova_results_list = kgroup_result,
  iv_name            = "Decade",
  dv_name            = "Box Office (millions)",
  group_labels       = c("1950s", "1970-80s", "2000s", "2010s"),
  KEY                = TRUE,
  table_title        = "Descriptive Statistics for Box Office by Decade"
)

# ANOVA source table
create_anova_source_table(
  anova_results_list = kgroup_result,
  KEY                = TRUE,
  table_title        = "ANOVA Source Table for Box Office by Decade"
)

# LSD pairwise comparisons (answer key)
lsd_pairwise_KEY(kgroup_result)
```

## 6. Generate APA Write-Up

`apa_kgroup_anova_writeup()` produces a write-up that includes the
omnibus test, post-hoc comparisons (if significant), and effect size
interpretation.

```{r}
#| label: writeup
#| eval: false

writeup <- apa_kgroup_anova_writeup(
  kgroup_results_list = kgroup_result,
  dv_name             = "box office revenue",
  iv_name             = "decade",
  group_labels        = c("1950s", "1970-80s", "2000s", "2010s"),
  hypothesis          = list(
    direction = "different",
    rh_text   = "box office revenue would differ across decades"
  ),
  posthoc_tests       = c("LSD", "Tukey's HSD")
)

# cat(writeup)
```

## 7. Hypothesis Support Evaluation

`create_rh_support_text()` generates text evaluating whether pairwise
results support the research hypothesis:

```{r}
#| label: rh-support
#| eval: false

create_rh_support_text(kgroup_result)
```

## Complete Lab Setup Example

```{r}
#| label: full-example
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# ── Analysis ─────────────────────────────────────────────
RH1_kgroup <- anova_kgroup_answers(
  data         = superman,
  dv           = "box_office_mil",
  iv           = "decade",
  group_labels = c("1950s", "1970-80s", "2000s", "2010s")
)

# ── Worksheet Output ─────────────────────────────────────
RH1_omnibus_KEY <- format_bg_anova_results(
  rh_name = "RH1",
  vars = c("Decade", "Box Office (millions)"),
  anova_results_list = RH1_kgroup,
  iv_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  KEY = TRUE
)

# ── APA Tables ───────────────────────────────────────────
RH1_desc_table <- create_bg_anova_table(
  anova_results_list = RH1_kgroup,
  iv_name = "Decade", dv_name = "Box Office (millions)",
  group_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  KEY = TRUE,
  table_title = "Descriptive Statistics for Box Office by Decade"
)

RH1_source_table <- create_anova_source_table(
  anova_results_list = RH1_kgroup, KEY = TRUE
)

# ── Write-Up ─────────────────────────────────────────────
RH1_writeup <- apa_kgroup_anova_writeup(
  kgroup_results_list = RH1_kgroup,
  dv_name = "box office revenue",
  iv_name = "decade",
  group_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  hypothesis = list(
    direction = "different",
    rh_text = "box office revenue would differ across decades"
  )
)
```

## Key Functions Reference

| Function | Purpose |
|----------|---------|
| `anova_kgroup_answers()` | K-group ANOVA with LSD/HSD and pairwise comparisons |
| `lsd_hsd_calculator()` | Compute LSD and HSD minimum mean differences |
| `pr_means_to_r()` | Pairwise r effect size from means and MSE |
| `create_rh_support_text()` | Hypothesis support evaluation text |
| `format_bg_anova_results()` | Markdown output for worksheets (KEY/BLANK) |
| `create_bg_anova_table()` | APA descriptives table (flextable) |
| `create_anova_source_table()` | ANOVA source table (flextable) |
| `lsd_pairwise_KEY()` | Pairwise comparison answer key |
| `apa_kgroup_anova_writeup()` | Full APA write-up with post-hoc results |
