---
title: "K-Group Chi-Square (k×2)"
subtitle: "Chi-square with 3+ groups and pairwise comparisons"
order: 6
---

## Overview

When one (or both) categorical variables has three or more levels, a
significant omnibus chi-square must be followed by pairwise 2×2 comparisons.
`chi_square_kgroup_answers()` handles the full workflow: the omnibus test,
all pairwise comparisons, and effect sizes.

## 1. Run the Analysis

```{r}
#| label: run-kgroup-chi
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# Run k-group chi-square
kgroup_chi <- chi_square_kgroup_answers(
  data       = superman,
  var1       = "decade",
  var2       = "tomatometer",
  var1_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels = c("Rotten", "Fresh")
)

# Omnibus results
kgroup_chi$Chi_Square

# Observed frequencies
kgroup_chi$Observed

# Pairwise comparisons
kgroup_chi$Pairwise
```

### What `chi_square_kgroup_answers()` returns

| Element | Contents |
|---------|----------|
| `$Chi_Square` | `chi_square`, `p_value`, `df`, `n` |
| `$Observed` | Matrix of observed cell frequencies |
| `$Expected` | Matrix of expected cell frequencies |
| `$Pairwise` | List of all pairwise 2×2 comparisons |
| `$var1_labels` | Display labels for the k-group variable |
| `$var2_labels` | Display labels for the 2-level variable |

### Pairwise comparison elements

Each entry in `$Pairwise` contains:

- `comparison` — which two groups are being compared
- `chi_sq`, `p_value`, `df` — test statistics for the 2×2 sub-table
- `effect_size` — phi/r coefficient for the pair
- `observed` — 2×2 observed frequency table

## 2. Create Checker Widgets

### Omnibus checker

`create_chisq_omnibus_table()` creates the omnibus statistics checker:

```{r}
#| label: omnibus-checker
#| eval: false

create_chisq_omnibus_table(
  rh_name          = "RH1",
  chi_results_list = kgroup_chi,
  var1_labels      = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels      = c("Rotten", "Fresh")
)
```

### Pairwise checker

`create_chisq_pairwise_checker()` creates fill-in-the-blank widgets
for each pairwise comparison:

```{r}
#| label: pairwise-checker
#| eval: false

create_chisq_pairwise_checker(
  rh_name          = "RH1",
  chi_results_list = kgroup_chi
)
```

## 3. Create APA Tables

### Combined results table

`create_chi_combined_table()` produces a single table with the omnibus
result and all pairwise comparisons:

```{r}
#| label: combined-table
#| eval: false

create_chi_combined_table(
  chi_results_list = kgroup_chi,
  var1_name        = "Decade",
  var2_name        = "Tomatometer",
  var1_labels      = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels      = c("Rotten", "Fresh"),
  table_title      = "Chi-Square Results for Decade × Tomatometer"
)
```

### Pairwise results table

```{r}
#| label: pairwise-table
#| eval: false

create_chisq_pairwise_table(
  chi_results_list = kgroup_chi,
  var1_labels      = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels      = c("Rotten", "Fresh")
)
```

## 4. Generate APA Write-Up

`apa_kgroup_chi_writeup()` produces a write-up covering the omnibus
test and any significant pairwise follow-ups:

```{r}
#| label: writeup
#| eval: false

writeup <- apa_kgroup_chi_writeup(
  chi_results_list = kgroup_chi,
  var1_name        = "decade",
  var2_name        = "Tomatometer rating",
  var1_labels      = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels      = c("Rotten", "Fresh"),
  hypothesis       = list(
    rh_text = "the distribution of Fresh vs Rotten ratings would differ across decades"
  ),
  include_posthoc  = TRUE
)

# cat(writeup)
```

## 5. Format Omnibus Results

`format_chi_omnibus_results()` creates markdown output for the omnibus
portion of the worksheet:

```{r}
#| label: format-omnibus
#| eval: false

format_chi_omnibus_results(
  rh_name          = "RH1",
  chi_results_list = kgroup_chi,
  KEY              = TRUE
)
```

## Complete Lab Setup Example

```{r}
#| label: full-example
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# ── Analysis ─────────────────────────────────────────────
RH1_kchi <- chi_square_kgroup_answers(
  data = superman,
  var1 = "decade",
  var2 = "tomatometer",
  var1_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels = c("Rotten", "Fresh")
)

# ── Checkers (for HTML) ──────────────────────────────────
# Omnibus checker
create_chisq_omnibus_table("RH1", RH1_kchi,
  var1_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels = c("Rotten", "Fresh")
)

# Pairwise checker
create_chisq_pairwise_checker("RH1", RH1_kchi)

# Pairwise answer key
chisq_pairwise_KEY(RH1_kchi)

# ── APA Tables ───────────────────────────────────────────
create_chi_combined_table(
  chi_results_list = RH1_kchi,
  var1_name = "Decade", var2_name = "Tomatometer",
  var1_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels = c("Rotten", "Fresh"),
  table_title = "Chi-Square Results for Decade × Tomatometer"
)

# ── Write-Up ─────────────────────────────────────────────
RH1_writeup <- apa_kgroup_chi_writeup(
  chi_results_list = RH1_kchi,
  var1_name = "decade",
  var2_name = "Tomatometer rating",
  var1_labels = c("1950s", "1970-80s", "2000s", "2010s"),
  var2_labels = c("Rotten", "Fresh"),
  hypothesis = list(
    rh_text = "the distribution of ratings would differ across decades"
  ),
  include_posthoc = TRUE
)
```

## Key Functions Reference

| Function | Purpose |
|----------|---------|
| `chi_square_kgroup_answers()` | K-group chi-square with pairwise comparisons |
| `create_chisq_omnibus_table()` | Omnibus statistics checker widget |
| `create_chisq_pairwise_checker()` | Pairwise comparison checker widget |
| `create_chisq_pairwise_table()` | Pairwise results table (flextable) |
| `create_chi_combined_table()` | Combined omnibus + pairwise table |
| `chisq_pairwise_KEY()` | Pairwise comparison answer key |
| `format_chi_omnibus_results()` | Markdown output for omnibus results |
| `apa_kgroup_chi_writeup()` | Full APA write-up with post-hoc |
