---
title: "Lab 7: Bivariate Analyses: Correlations and Chi-Squares"
format:
  hwchecker-html: default
learning_01: |
  Practice using SPSS to perform Pearson's correlation and Chi-Square analyses.
execute: 
  echo: false
---
```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: false

library(dplyr)
library(webexercises)
library(psych350data)
library(psych350lab)
library(glue)

data(superman, package = "psych350data")
superman <- prep_data(superman, "superman", keep_labels = TRUE)

data(hot_ones, package = "psych350data")
hot_ones <- prep_data(hot_ones, "hot_ones", keep_labels = TRUE)
```

::: instructions
## Instructions

For today's lab you will need to complete the `{{< meta title >}} Worksheet.`


### Learning Objectives

1.  {{< meta learning_01 >}}

### Download Data Files

[{{< fa file-arrow-down >}} Superman Data (.sav)](../datasets/superman/superman_data.sav){.btn .btn-primary download="superman_data.sav"}
[{{< fa file-arrow-down >}} Hot Ones Data (.sav)](../datasets/hotones/hot_ones_data.sav){.btn .btn-primary download="hot_ones_data.sav"}

:::

::: vspace
:::


:::: panel-tabset

::: {.content-hidden} 

### Today's Assignment

[{{< fa file-word title="Word document" >}} Download Lab](https://canvas.unl.edu/courses/205040/pages/lab-syllabus-and-schedule){.btn-primary .btn role="button"}

## Lab Slides

Either click on the slide area below or click [here](slides.qmd) to view it in fullscreen.

Use your keypad to navigate the slides.

::: {.container style="position: relative;overflow: hidden;width: 100%;min-height:600px;"}
<iframe src="slides.html" style="position=absolute;top: 0;left: 0;bottom: 0;right: 0;width: 100%;min-height:600px;">

</iframe>
:::

:::

## Homework checker
```{r}
#| label: practice-checker
#| results: asis
#| echo: false

library(glue)

# SAMPLE QUESTION
cat("\n**How to use the Homework Checker!**\n\n")

cat("\n\n")
cat("1. Type your answer into each blank box. Round to 2 decimal places where appropriate. \n")
cat("2. Select your choice from the dropdown menu for interpretability\n")
cat("3. The box will turn **green** if your answer is correct\n")
cat("4. The box will turn **red** if your answer is incorrect - try again!\n")
cat("5. Do not forget to take a screenshot of the correct answers to include on your lab worksheet!\n\n")


```


```{r}
#| label: create-tables
#| results: asis
#| echo: false

###########################################
# WALKTHROUGH SECTION CORR
###########################################

superman_data <- superman |>
  mutate(lbd_likes_thou = ldb_likes / 1000)

RH1 <- corr_answers(
  data = superman_data,
  var1 = "height_diff",
  var2 = "lbd_likes_thou"
)

RH1_vars <- c("height_diff", "lbd_likes_thou")
RH1_table <- create_corr_checker("RH1", RH1_vars, RH1)

cat("\n## **Walkthrough Answers**\n\n")
cat("\n### Enter the descriptives and Pearson's correlation statistics for RH1:\n\n")
RH1_table

###########################################
# WALKTHROUGH SECTION CHISQ
###########################################

# After prep_data(): clark_grp is 1 = Under 6ft, 2 = 6ft or taller
#                    tomatometer is 1 = Rotten, 2 = Fresh
# Numeric ordering: clark_grp rows = 1, 2; tomatometer cols = 1, 2
RH2 <- chi_square_answers(
  data = superman_data,
  var1 = "clark_grp",
  var2 = "tomatometer"
)

RH2_table <- create_chisq_checker(
  rh_name = "RH2",
  chi_results_list = RH2,
  var1_labels = c("Under 6ft", "6ft or taller"),
  var2_labels = c("Rotten", "Fresh")
)

cat("\n### Enter the frequencies and Pearson's chi-square statistics for RH2:\n\n")
RH2_table

###########################################
# YOUR TURN SECTION CORR
###########################################

# After prep_data(): occupation is numeric, result is 1 = Succeeded
hotones_data <- hot_ones |>
  mutate(
    total_SHU_mil = (SHU_1 + SHU_2 + SHU_3 + SHU_4 + SHU_5 +
                    SHU_6 + SHU_7 + SHU_8 + SHU_9 + SHU_10) / 1000000,
    engagement = comments / views,
    actor = case_when(
      is.na(occupation) ~ NA_real_,
      occupation %in% c(3, 4, 7) ~ 1,
      TRUE ~ 0
    ),
    popular = if_else(likes <= 125000, 1, 2),
    age = if_else(age == -99, NA_real_, age),
    age_grp = if_else(age < 35, 1, 2)
  ) |>
  filter(result == 1)

RH3 <- corr_answers(
  data = hotones_data,
  var1 = "total_SHU_mil",
  var2 = "views"
)

RH3_vars <- c("total_SHU_mil", "views")
RH3_table <- create_corr_checker("RH3", RH3_vars, RH3)

cat("\n## **Your Turn Answers**\n\n")
cat("\n### Enter the descriptives and Pearson's correlation statistics for RH3:\n\n")
RH3_table

RH4 <- corr_answers(
  data = hotones_data,
  var1 = "total_SHU_mil",
  var2 = "engagement"
)

RH4_vars <- c("total_SHU_mil", "engagement")
RH4_table <- create_corr_checker("RH4", RH4_vars, RH4)

cat("\n### Enter the descriptives and Pearson's correlation statistics for RH4:\n\n")
RH4_table

###########################################
# YOUR TURN SECTION CHISQ
###########################################

# actor: 0 = Non-Actor, 1 = Actor (numeric ordering: 0, 1)
# popular: 1 = Less Popular, 2 = More Popular (numeric ordering: 1, 2)
RH5 <- chi_square_answers(
  data = hotones_data,
  var1 = "actor",
  var2 = "popular"
)

RH5_table <- create_chisq_checker(
  rh_name = "RH5",
  chi_results_list = RH5,
  var1_labels = c("Non-Actor", "Actor"),
  var2_labels = c("Less Popular (≤125k)", "More Popular (>125k)")
)

cat("\n### Enter the frequencies and Pearson's chi-square statistics for RH5:\n\n")
RH5_table

hotones_rh6 <- hotones_data |>
  filter(!is.na(age) & !is.na(likes))

RH6 <- chi_square_answers(
  data = hotones_rh6,
  var1 = "age_grp",
  var2 = "popular"
)

RH6_table <- create_chisq_checker(
  rh_name = "RH6",
  chi_results_list = RH6,
  var1_labels = c("Under 35", "35 and Over"),
  var2_labels = c("Less Popular (≤125k)", "More Popular (>125k)")
)

cat("\n### Enter the frequencies and Pearson's chi-square statistics for RH6:\n\n")
RH6_table
```
```


::::

