---
title: "One-Way ANOVA (BG & WG)"
subtitle: "Between-groups and within-groups one-way ANOVA for two groups"
order: 4
---

## Overview

This guide covers two-group one-way ANOVA in **psych350lab**:

- **Between-groups (BG)** — comparing means across two independent groups
- **Within-groups (WG)** — comparing means across two repeated-measures
  conditions

Both follow the same workflow: run the analysis, format results for
worksheets, create APA tables, generate write-ups, and build homework
checkers.

## Between-Groups ANOVA

### 1. Run the Analysis

`bg_anova_answers()` performs a one-way between-groups ANOVA with
descriptive statistics for each group.

```{r}
#| label: run-bg
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# Run between-groups ANOVA
bg_result <- bg_anova_answers(
  data = superman,
  iv   = "clark_grp",
  dv   = "rt_audience_score"
)

# Inspect the results
bg_result$ANOVA
bg_result$Descriptives
```

#### What `bg_anova_answers()` returns

| Element | Contents |
|---------|----------|
| `$ANOVA` | `F`, `p_value`, `df_between`, `df_within`, `mse` |
| `$Descriptives` | Tibble with `iv`, `mean`, `sd`, `n`, `sem` per group |
| `$Sample_Size` | Total valid cases |

### 2. Format Results for Worksheets

`format_bg_anova_results()` creates markdown output for lab worksheets
with KEY/BLANK toggle.

```{r}
#| label: format-bg
#| eval: false

# Answer key
bg_KEY <- format_bg_anova_results(
  rh_name            = "RH1",
  vars               = c("Height Group", "Audience Score"),
  anova_results_list = bg_result,
  iv_labels          = c("Under 6ft", "6ft or taller"),
  KEY                = TRUE
)

# Student worksheet
bg_BLANK <- format_bg_anova_results(
  rh_name            = "RH1",
  vars               = c("Height Group", "Audience Score"),
  anova_results_list = bg_result,
  iv_labels          = c("Under 6ft", "6ft or taller"),
  KEY                = FALSE
)

# cat(bg_KEY)
```

### 3. Create APA Tables

```{r}
#| label: bg-tables
#| eval: false

# Descriptives table
create_bg_anova_table(
  anova_results_list = bg_result,
  iv_name       = "Height Group",
  dv_name       = "Audience Score",
  group_labels  = c("Under 6ft", "6ft or taller"),
  KEY           = TRUE,
  table_title   = "Descriptive Statistics for Audience Score by Height Group"
)

# ANOVA source table (SPSS-style)
create_anova_source_table(
  anova_results_list = bg_result,
  KEY                = TRUE,
  table_title        = "ANOVA Source Table"
)
```

### 4. Generate Write-Up

```{r}
#| label: bg-writeup
#| eval: false

writeup <- apa_bg_anova_writeup(
  anova_results_list = bg_result,
  iv_name            = "height group",
  dv_name            = "audience score",
  group_labels       = c("under 6ft", "6ft or taller"),
  hypothesis         = list(
    direction = "greater",
    rh_text   = "taller actors' movies would receive higher audience scores"
  )
)

# cat(writeup)
```

### 5. Pairwise Effect Size

`pr_means_to_r()` computes r from two group means using the omnibus MSE:

```{r}
#| label: effect-size
#| eval: false

pr_means_to_r(
  mean1 = bg_result$Descriptives$mean[1],
  mean2 = bg_result$Descriptives$mean[2],
  mse   = bg_result$ANOVA$mse,
  n1    = bg_result$Descriptives$n[1],
  n2    = bg_result$Descriptives$n[2]
)
```

## Within-Groups ANOVA

### 1. Run the Analysis

`wg_anova_answers()` performs a paired/repeated-measures ANOVA for two
conditions. Each row in the data represents one participant measured
twice.

```{r}
#| label: run-wg
#| eval: false

# Example: comparing critics score vs audience score for each movie
wg_result <- wg_anova_answers(
  data      = superman,
  dv1       = "rt_critics_score",
  dv2       = "rt_audience_score",
  dv1_label = "Critics Score",
  dv2_label = "Audience Score"
)

wg_result$ANOVA
wg_result$Descriptives
```

#### What `wg_anova_answers()` returns

| Element | Contents |
|---------|----------|
| `$ANOVA` | `F`, `p_value`, `df_effect`, `df_error`, `mse` |
| `$Descriptives` | Tibble with `condition`, `condition_label`, `mean`, `sd`, `n`, `sem` |
| `$Sample_Size` | Number of complete cases |

### 2. Format Results for Worksheets

```{r}
#| label: format-wg
#| eval: false

wg_KEY <- format_wg_anova_results(
  rh_name            = "RH1",
  dv_name            = "Rating Score",
  anova_results_list = wg_result,
  condition_labels   = c("Critics Score", "Audience Score"),
  KEY                = TRUE
)

# cat(wg_KEY)
```

### 3. Create APA Tables

```{r}
#| label: wg-tables
#| eval: false

create_wg_anova_table(
  anova_results_list = wg_result,
  dv_name            = "Rating Score",
  condition_labels   = c("Critics Score", "Audience Score"),
  KEY                = TRUE,
  table_title        = "Descriptive Statistics for Rating Scores"
)
```

### 4. Generate Write-Up

```{r}
#| label: wg-writeup
#| eval: false

writeup <- apa_wg_anova_writeup(
  anova_results_list = wg_result,
  condition_labels   = c("critics score", "audience score"),
  dv_name            = "rating score",
  hypothesis         = list(
    direction = "greater_dv1",
    rh_text   = "critics scores would be higher than audience scores"
  )
)

# cat(writeup)
```

## Inline APA Statistics

Both BG and WG results work with `apa_inline_anova()`:

```{r}
#| label: inline
#| eval: false

# BG: *F*(1, 28) = 3.45, *MSE* = 125.67, *p* = .073
apa_inline_anova(bg_result)

# WG: *F*(1, 29) = 8.12, *MSE* = 98.34, *p* = .008
apa_inline_anova(wg_result)

# Without MSE
apa_inline_anova(bg_result, include_mse = FALSE)
```

## Key Functions Reference

### Between-Groups

| Function | Purpose |
|----------|---------|
| `bg_anova_answers()` | Run BG ANOVA, get F, p, descriptives |
| `format_bg_anova_results()` | Markdown output for worksheets (KEY/BLANK) |
| `create_bg_anova_table()` | APA descriptives table (flextable) |
| `create_anova_source_table()` | ANOVA source table (flextable) |
| `apa_bg_anova_writeup()` | Full APA write-up |
| `pr_means_to_r()` | Pairwise r effect size from means |

### Within-Groups

| Function | Purpose |
|----------|---------|
| `wg_anova_answers()` | Run WG ANOVA, get F, p, descriptives |
| `format_wg_anova_results()` | Markdown output for worksheets (KEY/BLANK) |
| `create_wg_anova_table()` | APA descriptives table (flextable) |
| `apa_wg_anova_writeup()` | Full APA write-up |

### Shared

| Function | Purpose |
|----------|---------|
| `apa_inline_anova()` | Inline APA-formatted F statistic |
| `create_anova_source_table()` | ANOVA source table (works with both) |
