---
title: "Multiple Linear Regression"
subtitle: "Multiple predictors with univariate, bivariate, and multivariate results"
order: 8
---

## Overview

`linear_reg_answers()` runs a complete multiple linear regression analysis
that mirrors the SPSS workflow: univariate descriptive statistics for each
variable, bivariate correlations between all pairs, and the multivariate
regression model with standardized and unstandardized coefficients.

## 1. Run the Analysis

```{r}
#| label: run-regression
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# Multiple regression with mixed predictor types
reg_result <- linear_reg_answers(
  data              = superman,
  criterion         = "rt_audience_score",
  quant_predictors  = c("clark_height_in", "box_office_mil"),
  binary_predictors = c("tomatometer"),
  quant_labels      = c("Actor Height (in)", "Box Office (millions)"),
  binary_labels     = c("Tomatometer (Fresh vs Rotten)"),
  criterion_label   = "Audience Score"
)

# Inspect the structure
str(reg_result, max.level = 2)
```

### What `linear_reg_answers()` returns

| Element | Contents |
|---------|----------|
| `$Univariate` | Descriptives (M, SD, n) for all variables |
| `$Bivariate` | Full correlation matrix for all variables |
| `$Model` | `R`, `R_sq`, `adj_R_sq`, `F`, `p_value`, `df_regression`, `df_residual` |
| `$Predictors` | Tibble with `b`, `SE`, `beta`, `t`, `p_value`, `sr2` for each predictor |
| `$Correlations` | Bivariate r for each predictor with the criterion |

### Arguments explained

- **`criterion`** — the dependent variable (Y)
- **`quant_predictors`** — continuous predictor variables
- **`binary_predictors`** — dichotomous predictors (coded 0/1 or 1/2)
- **`*_labels`** — display labels for tables and write-ups
- **`verbose`** — if TRUE, prints intermediate results

## 2. Model Statistics Checker

`create_regression_model_checker()` creates an interactive widget for
students to enter model-level statistics (R, R², F, p):

```{r}
#| label: model-checker
#| eval: false

create_regression_model_checker(reg_result)
```

## 3. Predictor Weight Checker

`create_regression_predictor_checker()` creates a widget for each
predictor's coefficients (b, SE, β, t, p, sr²):

```{r}
#| label: predictor-checker
#| eval: false

create_regression_predictor_checker(
  reg_results_list = reg_result,
  predictor_labels = c("Actor Height (in)", "Box Office (millions)",
                       "Tomatometer (Fresh vs Rotten)")
)
```

## 4. Interpretation Widgets

### Bivariate correlation interpretations

Students categorize each predictor-criterion correlation:

```{r}
#| label: corr-interp
#| eval: false

create_correlation_interpretations(
  reg_results_list = reg_result,
  output_format    = "table"
)
```

### Regression weight interpretations

Students categorize each predictor's contribution to the model:

```{r}
#| label: weight-interp
#| eval: false

create_regression_weight_interpretations(
  reg_results_list = reg_result,
  output_format    = "table"
)
```

## 5. Model Statistics Display

These functions produce formatted displays of model-level results:

```{r}
#| label: model-display
#| eval: false

# Model statistics (R, R², F, p)
linear_reg_model_statistics(reg_result)

# Model evaluation criteria
linear_reg_model_evaluation(reg_result)

# Category legend for interpretation
regression_category_legend()
```

## 6. Combined Results Table

`create_regression_combined_table()` produces a comprehensive flextable
with all regression results:

```{r}
#| label: combined-table
#| eval: false

create_regression_combined_table(reg_result)
```

## 7. Interpretation Flextables

These produce formatted flextables for the interpretation sections:

```{r}
#| label: interp-tables
#| eval: false

# Bivariate correlation interpretation table
create_correlation_interp_table(reg_result)

# Regression weight interpretation table
create_regwt_interp_table(reg_result)
```

## Complete Lab Setup Example

```{r}
#| label: full-example
#| eval: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")

# ── Analysis ─────────────────────────────────────────────
RH1_reg <- linear_reg_answers(
  data = superman,
  criterion = "rt_audience_score",
  quant_predictors = c("clark_height_in", "box_office_mil"),
  binary_predictors = c("tomatometer"),
  quant_labels = c("Actor Height (in)", "Box Office (millions)"),
  binary_labels = c("Tomatometer"),
  criterion_label = "Audience Score"
)

# ── Checkers (for HTML) ──────────────────────────────────
# Model statistics checker
create_regression_model_checker(RH1_reg)

# Predictor weight checker
create_regression_predictor_checker(
  RH1_reg,
  predictor_labels = c("Actor Height", "Box Office", "Tomatometer")
)

# Interpretation widgets
create_correlation_interpretations(RH1_reg, output_format = "table")
create_regression_weight_interpretations(RH1_reg, output_format = "table")

# ── APA Tables ───────────────────────────────────────────
# Combined regression results table
create_regression_combined_table(RH1_reg)

# Model statistics display
linear_reg_model_statistics(RH1_reg)
linear_reg_model_evaluation(RH1_reg)

# Category legend
regression_category_legend()

# Interpretation tables
create_correlation_interp_table(RH1_reg)
create_regwt_interp_table(RH1_reg)
```

## Key Functions Reference

### Analysis

| Function | Purpose |
|----------|---------|
| `linear_reg_answers()` | Run multiple regression with full results |

### Interactive Checkers

| Function | Purpose |
|----------|---------|
| `create_regression_model_checker()` | Model statistics checker (R, R², F, p) |
| `create_regression_predictor_checker()` | Predictor coefficients checker |
| `create_correlation_interpretations()` | Bivariate correlation categorization |
| `create_regression_weight_interpretations()` | Regression weight categorization |

### Display Tables

| Function | Purpose |
|----------|---------|
| `create_regression_combined_table()` | Combined results table (flextable) |
| `linear_reg_model_statistics()` | Model R, R², F, p display |
| `linear_reg_model_evaluation()` | Model evaluation criteria |
| `regression_category_legend()` | Interpretation category legend |
| `create_correlation_interp_table()` | Correlation interpretation table |
| `create_regwt_interp_table()` | Regression weight interpretation table |

### Aliases

Several functions have shorter aliases for convenience:

| Full name | Alias |
|-----------|-------|
| `linear_reg_model_statistics()` | `regression_model_statistics()` |
| `linear_reg_model_evaluation()` | `regression_model_evaluation()` |
| `create_linear_reg_combined_table()` | `create_regression_combined_table()` |
| `linear_reg_table_legend()` | `regression_table_legend()` |
| `linear_reg_category_legend()` | `regression_category_legend()` |
