---
title: "Lab 06"
subtitle: Data Entry, Transformations & Univariate Stats
params: 
  hide_answers: false
  blank: true
format: 
  docx:
    reference-doc: lab-template.docx
---
```{r}
#| label: prep-data
#| echo: false
#| warning: false
#| message: false

library(psych350lab)
library(psych350data)
library(dplyr)

data(superman, package = "psych350data")
#superman <- prep_data(superman, "superman")
# Use keep_labels so they are automatically detected and added to table
superman <- prep_data(superman, "superman", keep_labels = TRUE)

# Walk-through: year > 1975
walkthrough_data <- superman |>
  filter(year > 1975)

walkthrough_vars <- c("num", "year", "clark_height", "lois_height",
                      "clark_height_in", "lois_height_in", "height_diff",
                      "type", "clark_grp", "height_gap")

summary_stats_walkthrough <- compute_summary_stats(
  walkthrough_data |> select(all_of(walkthrough_vars))
)

# Take-home: rt_critics_count >= 5
takehome_vars <- c("rt_critics_score", "rt_critics_count",
                   "rt_audience_score", "rt_audience_count",
                   "ldb_likes", "ldb_scores",
                   "rt_avg", "rt_diff", "tomatometer", "popular")

takehome_data <- superman |>
  filter(rt_critics_count >= 5)

summary_stats_takehome <- compute_summary_stats(
  takehome_data |> select(all_of(takehome_vars))
)
```


::: {custom-style="Box"}
[Instructions]{custom-style="section-title"}

This worksheet will walk us through how to enter data into SPSS, how to transform/recode that data as needed, and run some basic univariate statistics.

**Looking Forward:** These basic SPSS skills will be needed for all of the future SPSS assignments, especially the final project. It will be important to have a firm grasp of the recode/select/transform operations, as your final projects will require you to apply these in a unique way.

This worksheet will first walk us through how to enter data into SPSS, transform/recode that data as needed, and run some basic univariate statistics. 

:::

# Walk-through (10 points)

The dataset (shown below) includes the following variables (with values & labels for qualitative variables):

-   `num` -- unique number for each actor in dataset
-   `year` -- year of release
-   `type` -- type of media (1=Film, 2=TV Series, 3=Serial)
-   `clark_height` -- height of Clark Kent/Superman actor in meters (-99 means data missing)
-   `lois_height` -- height of Lois Lane actor in meters (-99 means data missing)

| *Note that the "-99" values mean that we will need to handle missing values.*

### Things to do:

1.  Using the **Variable View** specify the attributes of each variable (be sure to use Labels and Values).

2.  Using the **Data View** enter the data.

{{< pagebreak >}}
```{r}
#| label: data-entry-walkthrough
#| message: false
#| warning: false
#| echo: false

superman |>
  select(num, year, type, clark_height, lois_height) |>
  mutate(across(where(is.numeric), \(x) ifelse(is.na(x), -99, x))) |>
  flextable::flextable() |>
  flextable::colformat_double(j = "year", big.mark = "", digits = 0) |>
  flextable::theme_box() |>
  flextable::align(align = "center", part = "all") |>
  flextable::fontsize(size = 9, part = "all") |>
  flextable::width(j = "num", width = 0.6) |>
  flextable::width(j = "year", width = 1.1) |>
  flextable::width(j = "type", width = 1.1) |>
  flextable::width(j = c("clark_height", "lois_height"), width = 1.4)
```


**Do each of the following transformations  -- one at a time!!**

-   Be sure to check if you got what you want by looking at the variable view & running a frequencies
-   Be sure to make Labels and Values for each (mind-numbing at the time, but important later…)

3.  Make two new variables called `clark_height_in` and `lois_height_in` that reflect the height of each actor in inches rather than meters. Do this by computing the values of `clark_height` and `lois_height` from meters to inches. (*HINT: 1 meter = 39.37 inches!*)

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Convert Clark Kent height from meters to inches.
COMPUTE clark_height_in = clark_height * 39.37.
EXECUTE.

* Convert Lois Lane height from meters to inches.
COMPUTE lois_height_in = lois_height * 39.37.
EXECUTE.

* Add variable labels.
VARIABLE LABELS clark_height_in 'Clark Kent/Superman Actor Height (inches)'.
VARIABLE LABELS lois_height_in 'Lois Lane Actor Height (inches)'.
```
`r if (params$hide_answers) ":::"`

4.  Make a variable called `clark_grp` by recoding the `clark_height_in` variable into a new variable (we never want to recode into the original variable as it overwrites our data!). This new variable will group Clark Kent/Superman actors into those who are: `1 = under 6ft tall, 2 = over 6ft tall`. (*HINT: 1 foot = 12 inches*)

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Recode Clark height into groups (72 inches = 6 feet).
RECODE clark_height_in (Lowest thru 72=1) (72 thru Highest=2) INTO clark_grp.
EXECUTE.

* Add variable label.
VARIABLE LABELS clark_grp 'Clark Kent Height Group'.

* Add value labels.
VALUE LABELS clark_grp
  1 'Under 6ft'
  2 '6ft or taller'.
```
`r if (params$hide_answers) ":::"`

5.  Make a variable called `height_diff` that reflects the height difference between Clark Kent and Lois Lane actors in inches. Do this by computing the difference (`height_diff` = `clark_height_in` - `lois_height_in`).

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Compute height difference between Clark and Lois.
COMPUTE height_diff = clark_height_in - lois_height_in.
EXECUTE.

* Add variable label.
VARIABLE LABELS height_diff 'Height Difference (Clark - Lois) in inches'.
```
`r if (params$hide_answers) ":::"`

6.  Make a variable called `height_gap` by recoding `height_diff` into a new variable. This will categorize the height difference into three categories: `1 = minimal (< 6 inches), 2 = average (6-8 inches), and 3 = big (> 8 inches)`.

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Recode height difference into categories.
RECODE height_diff (Lowest thru 6=1) (6 thru 8=2) (8 thru Highest=3) INTO height_gap.
EXECUTE.

* Add variable label.
VARIABLE LABELS height_gap 'Height Gap Category'.

* Add value labels.
VALUE LABELS height_gap
  1 'Minimal (< 6 inches)'
  2 'Average (6-8 inches)'
  3 'Big (> 8 inches)'.
```

`r if (params$hide_answers) ":::"`

7.  Select only those cases where the actors debuted in media that was released after 1975 (using `year`) to include in your analyses.
(*HINT: After selecting cases, you should see some rows with a diagonal line through the row number in Data View, indicating they are not selected. The analyses you run will only include the selected cases.*)

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Select cases where year is greater than 1975.
USE ALL.
COMPUTE filter_$=(year > 1975).
VARIABLE LABELS filter_$ 'year > 1975 (FILTER)'.
VALUE LABELS filter_$ 0 'Not Selected' 1 'Selected'.
FORMATS filter_$ (f1.0).
FILTER BY filter_$.
EXECUTE.
```
`r if (params$hide_answers) ":::"`

8.  Get the mean, std, and SEM for all variables (entered and created) and enter those on the homework checker. The checker will also ask you whether the mean for each variable can be meaningfully interpreted.

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Get descriptives for quantitative variables.
DESCRIPTIVES VARIABLES=num year clark_height lois_height clark_height_in lois_height_in height_diff
  /STATISTICS=MEAN STDDEV SEMEAN MIN MAX.

* Get frequencies for categorical variables.
FREQUENCIES VARIABLES=type clark_grp height_gap
  /ORDER=ANALYSIS.
```
`r if (params$hide_answers) ":::"`

9.  Grab a screenshot of the homework checker screen, showing all of your input answers in green (correct). Paste it at the end of the take-home section. **Your TA will not grade the assignment without this.**

10.  Following the example on the handout, and using the template below, make a "Table 1" for the following variables: `height_diff`, `clark_grp`, and `height_gap`

`r if (params$blank) "::: {.content-hidden}"`
```{r}
#| label: apa-walkthrough
#| results: asis
#| echo: false
#| message: false
#| warning: false

cat("**Table 1.**\n\n")

create_descriptive_table(
  data        = walkthrough_data,
  continuous  = "height_diff",
  categorical = c("clark_grp", "height_gap"),
  KEY         = FALSE,
  footnote    = "Note. See APA manual beginning on p. 210 for additional examples."
)
```

`r if (params$blank) ":::"`

`r if (params$hide_answers) "::: {.content-hidden}"`
```{r}
#| label: walkthrough-ANSKEY
#| results: asis
#| echo: false

cat("### **Walk-through Answer Key**\n\n")
cat("*Note: These statistics are based on cases where year > 1975*\n\n")

create_answer_table(
  vars           = walkthrough_vars,
  stats_data     = summary_stats_walkthrough,
  label          = "num",
  quantitative   = c("year", "clark_height", "lois_height",
                     "clark_height_in", "lois_height_in", "height_diff"),
  binary         = "clark_grp",
  multi_category = c("type", "height_gap")
)
```

```{r}
#| label: apa-walkthrough-KEY
#| echo: false
#| results: asis

cat("**Table 1**\n\n")

create_descriptive_table(
  data        = walkthrough_data,
  continuous  = "height_diff",
  categorical = c("clark_grp", "height_gap"),
  stats_data  = summary_stats_walkthrough,
  var_labels  = c(
    height_diff = "Height Difference",
    clark_grp   = "Height Group",
    height_gap  = "Height Gap Category"
  )
)
```

`r if (params$hide_answers) ":::"`

11.  **Save the dataset (and syntax windows)! You will use it in the next section!** *NOTE: Before you do, return to the "Select Cases" window and click the "All Cases" radio button to return to analyzing all the data!*

{{< pagebreak >}}

# Take-Home (20 points)

::: {custom-style="Box"}
[Instructions]{custom-style="section-title"}

In this section, you will expand the dataset you created in the previous section by adding new ratings variables. You will use two publicly available sources of data: [Rotten Tomatoes®](https://www.rottentomatoes.com/about) (an entertainment review site) and [Letterboxd](https://letterboxd.com/) (social media site that lets users rate films). First, open your previously saved dataset, then add the following new variables:

:::

-   `rt_critics_score` -- Rotten Tomatoes critics score - represents the percentage of professional critic reviews that are positive for a given film or television show (0-100 scale; -99 means data missing)
-   `rt_critics_count` -- number of critic reviews on Rotten Tomatoes (-99 means data missing)
-   `rt_audience_score` -- Rotten Tomatoes audience score represents the percentage of audience reviews that are positive for a given film or television show (0-100 scale; -99 means data missing)
-   `audience_count` -- number of audience ratings number of critic reviews on Rotten Tomatoes (-99 means data missing)
-   `ldb_likes` -- total number of users that have "liked" the film on Letterboxd (-99 means data missing)
-   `ldb_scores` -- Letterboxd users average rating (1-5 stars; -99 means data missing)

| *Note that the "-99" values mean that we will need to handle missing values.*

### Things to do:

1.  Using the **Variable View** specify the attributes of each variable (be sure to use Labels and Values).

2.  Using the **Data View** enter the data.

{{< pagebreak >}}
```{r}
#| label: takehome-data
#| message: false
#| warning: false
#| echo: false

superman |>
  select(rt_critics_score, rt_audience_score, ldb_likes, ldb_scores) |>
  mutate(across(where(is.numeric), \(x) ifelse(is.na(x), -99, x))) |>
  flextable::flextable() |>
  flextable::theme_box() |>
  flextable::align(align = "left", part = "all") |>
  flextable::fontsize(size = 9, part = "all") |>
  flextable::autofit(add_w = 0)
```

-   **Do each of the following transformations  -- one at a time!!**
    -   Be sure to check if you got what you want by looking at the variable view & running a frequencies
    -   Be sure to make Labels and Values for each (mind-numbing at the time, but important later…)

3.  Make a new variable called `tomatometer` by recoding `rt_critics_score` using the Rotten Tomatoes' classification standards. We want this to represent whether the media was liked by more than 60% of critics. Let the categories be: `1 = Rotten; and 2 = Fresh`. 

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Recode critics score into Rotten/Fresh categories.
RECODE rt_critics_score (Lowest thru 60=1) (60 thru Highest=2) INTO tomatometer.
EXECUTE.

* Add variable label.
VARIABLE LABELS tomatometer 'Tomatometer Rating'.

* Add value labels.
VALUE LABELS tomatometer
  1 'Rotten'
  2 'Fresh'.
```
`r if (params$hide_answers) ":::"`

4.  Make a variable called `rt_avg` an average of the number of critics and audience members that liked the media on Rotten Tomatoes (using `rt_critics_score` and `rt_audience_score`). 

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Compute average of critics and audience scores.
COMPUTE rt_avg = (rt_critics_score + rt_audience_score) / 2.
EXECUTE.

* Add variable label.
VARIABLE LABELS rt_avg 'Rotten Tomatoes Average Score'.
```
`r if (params$hide_answers) ":::"`

5.  Make a variable called `rt_diff` that reflect the weighted difference between the number of critics and audience members that liked the media on Rotten Tomatoes. Do this by computing the difference: (`rt_critics_score × rt_critics_count) - rt_audience_score × rt_audience_count) / (rt_critics_count + rt_audience_count`).

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Compute weighted difference between critics and audience.
COMPUTE rt_diff = ((rt_critics_score * rt_critics_count) - (rt_audience_score * rt_audience_count)) / (rt_critics_count + rt_audience_count).
EXECUTE.

* Add variable label.
VARIABLE LABELS rt_diff 'Weighted Difference (Critics - Audience)'.
```
`r if (params$hide_answers) ":::"`

6.  Make a variable called `popular` by recoding `ldb_likes` into a new variable. This will categorize the popularity on Letterboxd into three categories: `1 = low (< 1,000 likes), 2 = mid (1,000-100,000 likes), 3 = high (> 100,000 likes)`.

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Recode Letterboxd likes into popularity categories.
RECODE ldb_likes (Lowest thru 1000=1) (1000 thru 100000=2) (100000 thru Highest=3) INTO popular.
EXECUTE.

* Add variable label.
VARIABLE LABELS popular 'Letterboxd Popularity'.

* Add value labels.
VALUE LABELS popular
  1 'Low (< 1,000 likes)'
  2 'Mid (1,000-100,000 likes)'
  3 'High (> 100,000 likes)'.
```
`r if (params$hide_answers) ":::"`

7.  Select only those cases where there are at least 5 critics scores (using `rt_critics_count`) to include in your analyses. 

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Select cases where there are at least 5 critics scores.
USE ALL.
COMPUTE filter_$=(rt_critics_count >= 5).
VARIABLE LABELS filter_$ 'rt_critics_count >= 5 (FILTER)'.
VALUE LABELS filter_$ 0 'Not Selected' 1 'Selected'.
FORMATS filter_$ (f1.0).
FILTER BY filter_$.
EXECUTE.
```
`r if (params$hide_answers) ":::"`

8.  Get the mean, std, and SEM for all the **new variables** (entered and created) and enter those on the homework checker. The checker will also ask you whether the mean, std, and sem for each variable can be meaningfully interpreted. 

`r if (params$hide_answers) "::: {.content-hidden}"`
```default
# SPSS SYNTAX

* Get descriptives for quantitative variables.
DESCRIPTIVES VARIABLES=rt_critics_score rt_critics_count rt_audience_score rt_audience_count ldb_likes ldb_scores rt_avg rt_diff
  /STATISTICS=MEAN STDDEV SEMEAN MIN MAX.

* Get frequencies for categorical variables.
FREQUENCIES VARIABLES=tomatometer popular
  /ORDER=ANALYSIS.
```
`r if (params$hide_answers) ":::"`

9.  Grab a screenshot of the homework checker screen, showing all of your input answers in green (correct). Paste it at the end of this section. **Your TA will not grade the assignment without this.**

10.  Following the example on the handout, and the template you used above, make a "Table 2" for the following variables on the next page: `tomatometer`, `rt_avg`, `rt_diff`, and `popular`

`r if (params$blank) "::: {.content-hidden}"`

```{r}
#| label: apa-takehome
#| results: asis
#| echo: false
#| message: false
#| warning: false

cat("**Table 2.**\n\n")

create_descriptive_table(
  data        = superman,
  continuous  = c("rt_avg", "rt_diff"),
  categorical = c("tomatometer", "popular"),
  KEY         = FALSE
)
```

`r if (params$blank) ":::"`

11. **Save the dataset (and syntax windows)! You will need this dataset next week!** *NOTE: Before you do, return to the "Select Cases" window and click the "All Cases" radio button to return to analyzing all the data!*

`r if (params$hide_answers) "::: {.content-hidden}"`
```{r}
#| label: takehome-ANSKEY
#| results: asis
#| echo: false

cat("### **Take-Home Answer Key**\n\n")
cat("*Note: These statistics are based on cases where rt_critics_count >= 5*\n\n")

create_answer_table(
  vars           = takehome_vars,
  stats_data     = summary_stats_takehome,
  quantitative   = c("rt_critics_score", "rt_critics_count",
                     "rt_audience_score", "rt_audience_count",
                     "ldb_likes", "ldb_scores", "rt_avg", "rt_diff"),
  binary         = "tomatometer",
  multi_category = "popular"
)
```

```{r}
#| label: apa-takehome-ANSKEY
#| results: asis
#| echo: false

cat("**Table 2**\n\n")

create_descriptive_table(
  data        = takehome_data,
  continuous  = c("rt_avg", "rt_diff"),
  categorical = c("tomatometer", "popular"),
  stats_data  = summary_stats_takehome,
  var_labels  = c(
    rt_avg      = "RT Average Score",
    rt_diff     = "RT Weighted Difference",
    tomatometer = "Tomatometer",
    popular     = "Letterboxd Popularity"
  )
)
```
`r if (params$hide_answers) ":::"`

\vspace{0.5cm}

<br>

\vspace{0.5cm}

## Homework Checkers

::: {custom-style="AnswerBox"}
`Paste Homework Checkers screenshots here`: 
:::
