---
title: "SPSS-Style HTML Tables"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{SPSS-Style HTML Tables}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
 results = "asis"
)
```

## Overview
The `psych350lab` package provides functions to create SPSS-style statistical tables as HTML output. These tables mimic the appearance of SPSS output and are ideal for:

- Quarto HTML documents
- R Markdown HTML output
- Teaching materials that reference SPSS
- Reports where familiar SPSS formatting is desired

## Getting Started

```{r load-package}
library(psych350lab)
data(superman)
```

## Descriptive Statistics

### Basic Descriptives

Use `spss_descriptives_html()` to create a descriptive statistics table:
```{r descriptives, eval=FALSE}
spss_descriptives_html(superman, c("clark_height_in", "rt_critics_score", "rt_audience_score"))
```

This produces a table with N, Mean, Std. Deviation, Std. Error, Minimum, and Maximum for each variable.

### Grouped Descriptives

For ANOVA-style grouped descriptives with confidence intervals:

```{r grouped-desc, eval=FALSE}
spss_descriptives_by_group_html(superman, dv = "rt_critics_score", iv = "clark_grp")
```

## T-Tests

### Independent Samples T-Test

```{r ttest-ind, eval=FALSE}
spss_ttest_independent_html(superman, dv = "rt_critics_score", iv = "clark_grp")
```

This produces a complete independent samples t-test table including:

- Levene's test for equality of variances
- T-test results for both equal variances assumed and not assumed
- Mean difference, standard error, and 95% confidence interval

### Paired Samples T-Test

```{r ttest-paired, eval=FALSE}
spss_ttest_paired_html(superman, var1 = "rt_critics_score", var2 = "rt_audience_score")
```

### One-Sample T-Test

```{r ttest-one, eval=FALSE}
spss_ttest_one_html(superman, variable = "clark_height_in", mu = 72)
```

## ANOVA

### One-Way ANOVA

You can create an ANOVA table from raw data:

```{r anova-raw, eval=FALSE}
spss_anova_html(superman, dv = "rt_critics_score", iv = "clark_grp")
```

Or from pre-computed results using `bg_anova_answers()`:

```{r anova-results, eval=FALSE}
anova_result <- bg_anova_answers(superman, iv = "clark_grp", dv = "rt_critics_score")
spss_anova_html(anova_result, dv = "rt_critics_score")
```

### Levene's Test

```{r levene, eval=FALSE}
spss_levene_html(superman, dv = "rt_critics_score", iv = "clark_grp")
```

## Chi-Square Analysis

### Chi-Square Test

From raw data:

```{r chisq-raw, eval=FALSE}
spss_chisq_html(superman, var1 = "clark_grp", var2 = "tomatometer")
```

Or from pre-computed results:

```{r chisq-results, eval=FALSE}
chi_result <- chi_square_answers(superman, "clark_grp", "tomatometer")
spss_chisq_html(chi_result)
```

### Crosstabulation

```{r crosstab, eval=FALSE}
spss_crosstab_html(superman, var1 = "clark_grp", var2 = "tomatometer")
```

## Correlation Analysis

Create a correlation matrix with significance stars:

```{r correlations, eval=FALSE}
spss_correlations_html(
 superman,
 variables = c("clark_height_in", "rt_critics_score", "rt_audience_score")
)
```

The table includes:

- Pearson correlation coefficients
- Significance levels (2-tailed)
- Sample sizes
- Stars indicating significance (* p < .05, ** p < .01)

## Regression Analysis

For regression tables, first compute the results using `regression_answers()`:

```{r regression-setup}
# Prepare data (remove missing values)
sm <- superman[complete.cases(superman[, c("rt_critics_score", "clark_height_in", "rt_audience_score")]), ]

# Compute regression
reg_result <- regression_answers(
 data = sm,
 criterion = "rt_critics_score",
 quant_predictors = c("clark_height_in", "rt_audience_score"),
 quant_labels = c("Clark Height", "Audience Score"),
 criterion_label = "Critics Score"
)
```

### Model Summary

```{r reg-model, eval=FALSE}
spss_regression_model_html(reg_result)
```

### Regression ANOVA

```{r reg-anova, eval=FALSE}
spss_regression_anova_html(reg_result)
```

### Coefficients Table

```{r reg-coef, eval=FALSE}
spss_regression_coefficients_html(reg_result)
```

## Saving Tables as PNG

For presentations or documents that require image files, use `spss_descriptives_png()`:

```{r png-save, eval=FALSE}
spss_descriptives_png(
 superman,
 variables = c("clark_height_in", "rt_critics_score"),
 filename = "descriptives.png",
 width = 7,
 dpi = 150
)
```

Note: This requires the `gt` package to be installed.

## Using with Quarto

In Quarto documents, use the `results: asis` chunk option to render the HTML tables properly:

````markdown
```{r}
#| results: asis
library(psych350lab)
data(superman)
spss_descriptives_html(superman, c("clark_height_in", "rt_critics_score"))
```
````

## Integration with Package Calculation Functions

The HTML table functions integrate seamlessly with the package's calculation functions:

| Calculation Function | HTML Table Function |
|---------------------|---------------------|
| `chi_square_answers()` | `spss_chisq_html()`, `spss_crosstab_html()` |
| `bg_anova_answers()` | `spss_anova_html()` |
| `regression_answers()` | `spss_regression_model_html()`, `spss_regression_anova_html()`, `spss_regression_coefficients_html()` |

This allows you to compute statistics once and use the results for both interactive homework checkers and publication-ready tables.
