---
title: "SPSS-Style Typst Tables for PDF Output"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{SPSS-Style Typst Tables for PDF Output}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
 eval = FALSE
)
```

## Overview

The `psych350lab` package provides functions to create SPSS-style statistical tables as Typst markup. These are designed for use in Quarto documents with `format: typst` to produce high-quality PDF output that mimics SPSS table appearance.

## Prerequisites

### Quarto Setup

To use these functions, you need:

1. Quarto installed (version 1.3+)
2. A Quarto document with `format: typst`
3. A Typst template that defines `#spss_table()` and `#spss_table_complex()` functions

### Document YAML

Your Quarto document should have a YAML header like:

```yaml
---
title: "My Analysis"
format:
 typst:
   template: spss-template.typ
---
```

## Basic Usage

In Quarto chunks, use `results: asis` to output the raw Typst markup:

````markdown
```{r}
#| results: asis
library(psych350lab)
data(superman)
typst_descriptives(superman, c("clark_height_in", "rt_critics_score"))
```
````

## Descriptive Statistics

### Basic Descriptives

```{r}
typst_descriptives(superman, c("clark_height_in", "rt_critics_score", "rt_audience_score"))
```

### Grouped Descriptives (for ANOVA)

```{r}
typst_descriptives_by_group(superman, dv = "rt_critics_score", iv = "clark_grp")
```

## T-Tests

### Group Statistics

```{r}
typst_group_stats(superman, dv = "rt_critics_score", iv = "clark_grp")
```

### Independent Samples T-Test

```{r}
typst_ttest_independent(superman, dv = "rt_critics_score", iv = "clark_grp")
```

### Paired Samples T-Test

```{r}
typst_ttest_paired(superman, var1 = "rt_critics_score", var2 = "rt_audience_score")
```
### One-Sample T-Test

```{r}
typst_ttest_one(superman, variable = "clark_height_in", mu = 72)
```

## ANOVA

### One-Way ANOVA

```{r}
typst_anova_oneway(superman, dv = "rt_critics_score", iv = "clark_grp")
```

### Levene's Test

```{r}
typst_levene(superman, dv = "rt_critics_score", iv = "clark_grp")
```

### Post-Hoc Comparisons

Generate Tukey HSD and/or LSD post-hoc tables:

```{r}
# Both Tukey and LSD
typst_posthoc(superman, dv = "rt_critics_score", iv = "type", methods = c("tukey", "lsd"))

# Tukey only
typst_posthoc(superman, dv = "rt_critics_score", iv = "type", methods = "tukey")

# LSD only
typst_posthoc(superman, dv = "rt_critics_score", iv = "type", methods = "lsd")
```

## Chi-Square Analysis

### Chi-Square Test

```{r}
typst_chisq(superman, var1 = "clark_grp", var2 = "tomatometer")
```

The output includes:

- Pearson Chi-Square
- Continuity Correction (for 2x2 tables)
- N of Valid Cases
- Footnotes about expected counts

### Crosstabulation

```{r}
typst_crosstab(superman, var1 = "clark_grp", var2 = "tomatometer")
```

## Correlation Analysis

```{r}
typst_correlations(
 superman,
 variables = c("clark_height_in", "rt_critics_score", "rt_audience_score")
)
```

## Regression Analysis

The `typst_regression()` function outputs all three standard regression tables:

1. Model Summary (R, R², Adjusted R², Std. Error)
2. ANOVA table
3. Coefficients table

```{r}
typst_regression(superman, "rt_critics_score ~ clark_height_in + rt_audience_score")
```

## Non-Parametric Tests

### Kruskal-Wallis Test

```{r}
typst_kruskal(superman, dv = "rt_critics_score", iv = "type")
```

## Styled Text Boxes

Use `typst_box()` to create styled callout boxes for writeups, tips, warnings, etc.:

```{r}
# APA-style writeup box
typst_box("The results indicated a significant relationship...", "writeup")

# Tip in the margin
typst_box("Remember to check assumptions!", "tip", side = "right")

# Warning box
typst_box("Assumptions were violated.", "warning", side = "left")
```

Available box types:

- `"writeup"` - For APA-style results paragraphs
- `"tip"` - Helpful hints (margin note)
- `"warning"` - Cautions and warnings (margin note)
- `"important"` - Key information (margin note)
- `"note"` - General notes (margin note)

## Working with SPSS Data Files

When working with `.sav` files imported via `haven::read_sav()`, you can pass the original data to preserve value labels:

```{r}
library(haven)

# Read SPSS file
data_raw <- read_sav("mydata.sav")
data <- zap_labels(zap_missing(data_raw))

# Tables will use value labels from original data
typst_descriptives_by_group(data, dv = "score", iv = "group", data_raw = data_raw)
typst_crosstab(data, var1 = "gender", var2 = "treatment", data_raw = data_raw)
```

## Complete Example

Here's a complete Quarto document example:

````markdown
---
title: "Superman Analysis"
format:
 typst:
   template: spss-template.typ
---

```{r}
#| include: false
library(psych350lab)
data(superman)
```

## Descriptive Statistics

```{r}
#| results: asis
typst_descriptives(superman, c("clark_height_in", "rt_critics_score"))
```

## Group Comparison

```{r}
#| results: asis
typst_descriptives_by_group(superman, "rt_critics_score", "clark_grp")
```

```{r}
#| results: asis
typst_levene(superman, "rt_critics_score", "clark_grp")
```

```{r}
#| results: asis
typst_anova_oneway(superman, "rt_critics_score", "clark_grp")
```

## Interpretation

```{r}
#| results: asis
typst_box(
 "A one-way ANOVA revealed a significant effect of Clark height group on critics scores...",
 "writeup"
)
```
````

## Typst Template Requirements

Your Typst template needs to define these functions:

### `#spss_table()`

For simple tables with uniform column headers.

Parameters:

- `title`: Table title
- `subtitle`: Optional subtitle
- `columns`: Array of column headers
- `rows`: Array of row data
- `label_cols`: Number of label columns (default: 1)
- `footnotes`: Optional array of footnotes

### `#spss_table_complex()`

For tables with multi-level headers.

Parameters:

- `title`: Table title
- `subtitle`: Optional subtitle
- `ncols`: Number of columns
- `header_rows`: Array of header row arrays
- `rows`: Array of row data
- `label_cols`: Number of label columns (default: 1)
- `footnotes`: Optional array of footnotes

See the package documentation for example Typst template implementations.
