---
title: "SPSS-Style Plots"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{SPSS-Style Plots}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
 fig.width = 6,
 fig.height = 4.5
)
```

## Overview

The `psych350lab` package provides functions to create plots that mimic the appearance of SPSS charts. This includes:

- A complete ggplot2 theme (`theme_SPSS()`)
- SPSS color palettes (`palette_SPSS()`, `scale_color_SPSS()`, `scale_fill_SPSS()`)
- Convenience functions for common chart types (`spss_histogram()`, `spss_barchart()`)

## Getting Started

```{r load-package}
library(psych350lab)
library(ggplot2)
data(superman)
```

## SPSS Theme

### Basic Usage

Apply `theme_SPSS()` to any ggplot2 plot:

```{r theme-basic}
ggplot(superman, aes(x = clark_height_in, y = rt_critics_score)) +
 geom_point(color = "#1192E8", size = 3) +
 labs(x = "Clark Height (inches)", y = "Critics Score") +
 theme_SPSS()
```

### Modern vs Legacy Style

The theme supports two styles:

- **Modern** (default): White background, horizontal grid lines, no box border
- **Legacy**: Grey background, box border, no grid lines

```{r theme-legacy}
ggplot(superman, aes(x = clark_height_in, y = rt_critics_score)) +
 geom_point(color = "#3E58AC", size = 3) +
 labs(x = "Clark Height (inches)", y = "Critics Score") +
 theme_SPSS(version = "legacy")
```

### Discrete vs Continuous Scales

Adjust text sizing based on axis type:

```{r theme-discrete}
ggplot(superman, aes(x = factor(clark_grp), y = rt_critics_score)) +
 geom_boxplot(fill = "#1192E8", alpha = 0.7) +
 labs(x = "Height Group", y = "Critics Score") +
 theme_SPSS(scale.x = "discrete")
```

## SPSS Color Palettes

### Viewing the Palette

```{r palette-view}
# Get first 8 colors from modern palette
palette_SPSS(8)

# Get legacy palette colors
palette_SPSS(8, version = "legacy")
```

### Using with ggplot2

```{r palette-ggplot}
ggplot(superman, aes(x = rt_critics_score, y = rt_audience_score, color = factor(type))) +
 geom_point(size = 3) +
 scale_color_SPSS() +
 labs(x = "Critics Score", y = "Audience Score", color = "Type") +
 theme_SPSS()
```

### Fill Scale

```{r fill-scale}
ggplot(superman, aes(x = factor(type), fill = factor(type))) +
 geom_bar() +
 scale_fill_SPSS() +
 labs(x = "Superman Type", y = "Count") +
 theme_SPSS(scale.x = "discrete") +
 theme(legend.position = "none")
```

## SPSS-Style Histogram

The `spss_histogram()` function creates a histogram with:

- SPSS blue fill color
- Optional normal distribution overlay
- Mean, SD, and N annotation

```{r histogram-basic}
spss_histogram(superman, "clark_height_in")
```

### Customization Options

```{r histogram-custom}
spss_histogram(
 superman,
 variable = "rt_critics_score",
 bins = 15,
 show_normal = TRUE,
 show_stats = TRUE,
 fill_color = "#005D5D",
 version = "modern"
)
```

### Without Normal Curve

```{r histogram-no-normal}
spss_histogram(
 superman,
 variable = "rt_audience_score",
 show_normal = FALSE,
 show_stats = TRUE
)
```

### Legacy Style

```{r histogram-legacy}
spss_histogram(
 superman,
 variable = "clark_height_in",
 version = "legacy",
 fill_color = "#3E58AC"
)
```

## SPSS-Style Bar Chart

The `spss_barchart()` function creates frequency bar charts for categorical variables:

```{r barchart-basic}
spss_barchart(superman, "clark_grp")
```

### Customization

```{r barchart-custom}
spss_barchart(
 superman,
 variable = "type",
 show_counts = TRUE,
 fill_color = "#9F1853",
 version = "modern"
)
```

### Without Count Labels

```{r barchart-no-counts}
spss_barchart(
 superman,
 variable = "tomatometer",
 show_counts = FALSE
)
```

## Saving Plots

### Using spss_save_plot()

```{r save-plot, eval=FALSE}
# Create a plot
p <- spss_histogram(superman, "clark_height_in")

# Save as PNG
spss_save_plot(p, "histogram.png", width = 6, height = 4.5, dpi = 150)
```

### Using ggplot2's ggsave()

Since the functions return ggplot2 objects, you can also use `ggsave()`:

```{r ggsave, eval=FALSE}
p <- spss_histogram(superman, "clark_height_in")
ggsave("histogram.pdf", p, width = 6, height = 4.5)
```

## Building Custom Plots

You can combine `theme_SPSS()` and the color scales with any ggplot2 code:

### Scatter Plot with Regression Line

```{r custom-scatter}
ggplot(superman, aes(x = clark_height_in, y = rt_critics_score)) +
 geom_point(color = "#1192E8", size = 2, alpha = 0.7) +
 geom_smooth(method = "lm", color = "#FA4D56", se = TRUE, fill = "#FA4D56", alpha = 0.2) +
 labs(
   x = "Clark Height (inches)",
   y = "Critics Score",
   title = "Relationship Between Clark Height and Critics Score"
 ) +
 theme_SPSS()
```

### Grouped Box Plot

```{r custom-boxplot}
ggplot(superman, aes(x = factor(type), y = rt_critics_score, fill = factor(type))) +
 geom_boxplot(alpha = 0.8) +
 scale_fill_SPSS() +
 labs(x = "Superman Type", y = "Critics Score") +
 theme_SPSS(scale.x = "discrete") +
 theme(legend.position = "none")
```

### Faceted Plot

```{r custom-facet}
ggplot(superman, aes(x = rt_critics_score)) +
 geom_histogram(bins = 10, fill = "#1192E8", color = "black", linewidth = 0.3) +
 facet_wrap(~clark_grp) +
 labs(x = "Critics Score", y = "Frequency") +
 theme_SPSS()
```

## Number Formatting

Use `number_SPSS()` for axis labels without thousands separators (matching SPSS defaults):
```{r number-format}
# Create data with large numbers
big_data <- data.frame(
 x = c(1000, 5000, 10000, 25000),
 y = c(100, 200, 150, 300)
)

ggplot(big_data, aes(x = x, y = y)) +
 geom_point(size = 3, color = "#1192E8") +
 scale_x_continuous(labels = number_SPSS) +
 theme_SPSS()
```

## Theme Parameters Reference

| Parameter | Description | Default |
|-----------|-------------|---------|
| `base_size` | Base font size (pt) | 14 |
| `base_family` | Font family | "sans" |
| `base_line_size` | Line width | 0.75 |
| `base_rect_size` | Rectangle border width | 0.75 |
| `version` | "modern" or "legacy" | "modern" |
| `scales` | "continuous" or "discrete" | "continuous" |
| `scale.x` | Override for x-axis | Same as `scales` |
| `scale.y` | Override for y-axis | Same as `scales` |

## Color Palette Reference

### Modern Palette (SPSS v24+)

The first 10 colors:

1. `#1192E8` - Blue
2. `#005D5D` - Teal
3. `#9F1853` - Magenta
4. `#FA4D56` - Red
5. `#570408` - Dark Red
6. `#198038` - Green
7. `#002D9C` - Dark Blue
8. `#EE538B` - Pink
9. `#B28600` - Gold
10. `#009D9A` - Cyan

### Legacy Palette (SPSS <v24)

The first 10 colors:

1. `#3E58AC` - Blue
2. `#2EB848` - Green
3. `#D3CE97` - Tan
4. `#79287D` - Purple
5. `#FBF873` - Yellow
6. `#EF3338` - Red
7. `#48C2C5` - Cyan
8. `#CCCCCC` - Grey
9. `#7AAAD5` - Light Blue
10. `#0A562C` - Dark Green
